### 2025-03-07 analysis/paper/paper.qmd
work on discussion section and add references

---
title: "Indigenous Subsistence and Colonial Encounters: A Zooarchaeological Perspective from Kiwulan in Northeastern Taiwan"
author:
  - Li-Ying Wang:
      correspondence: "yes"
      email: liyingw15@gmail.com
      orcid: 0000-0001-9060-2125
      institute: Academia Sinica
title-block-published: "Last updated"
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Colonial encounters between Europeans and Indigenous peoples often had profound impacts on Indigenous societies and cultures in many parts of the world. In northeastern Taiwan during the Late Iron Age, local Indigenous groups experienced intense foriegn contact through regional trade networks in East Asia, which expanded further with the arrival of Europeans in the 17th century. This study focuses on local subsistence through faunal remains to explore how Indigenous communities responded to trade activities driven by Europeans and Han Chinese, particularly the high demand for natural resources such as deer. By analyzing faunal remains from the upper component of Kiwulan (AD 1350-1850), a late Iron Age site in northeastern Taiwan, this research dicusses the Indigenous participation in the deerskin trade and their adaptive strategies. The results suggest cultural persistence in deer consumption, along with increased reliance on suids during the European period as demand for deer products intensified. This paper also provides insights into colonial experiences from Indigenous settlements at the periphery of European colonial territories, a perspective rarely explored.

keywords: |
  Zooarchaeology; Iron Age; Taiwan; sika deer; trade; European colonization
highlights: |
  Indigenous community in northeastern Taiwan engaged in regional deer trade were identified from faunal remains.
  The high freqiencies of deer bones at Kiwulan reveals continously cultural persistence in deer consumption.
  Changes in deer butchery patterns and cooking practices indicates an intensified hunting during the European colonization period.
  Increased reliance on suids during the European period suggests adaptive strategies to sustain local subsistence.
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The actual document text starts here: -->

# Introduction
<!-- introduction about European expansion -->
European expansion since the 15th century was closely tied to long-distance trade, exploration, and colonization, which often had profound impacts on Indigenous societies. The impacts also varied across regions depending on specific geographic and historical circumstances, resulting in different forms of European colonization and power relations [@Jordan2009]. Understanding the diverse and dynamic colonial processes has become a key focus in archaeology, because it provides critical insights into the complex consequences of European expansion [@Gosden2004]. In East Asia, European colonization was mostly driven by the pursuit of raw materials, natural resources, or luxury goods for commercial purposes [@Prakash1997]. The new exchange systems and colonial institutions introduced by Europeans profoundly influenced local economies, social structures, or settlement patterns [@Christie1996; @Junker2004; @Lape2002]. However, recent studies emphasize that Indigenous groups were not just passively accept European colonization; instead, they actively responded through resistance, negotiation, and cultural persistence or resilience, leading to complex and entangled colonial experiences [@Torrence2003; @Silliman2005; @Mitchell2010; @Jordan2009; @Alexander1998; @Noe2023].

<!-- NE Taiwan as a case study -->
While many studies focus on prominent colonies, places on the periphery of colonial bases were also more and less impacted by European powers or associated trading systems [@Acabado2017; @Trabert2017]. Exploring these areas is important to gain a broader understanding of the varied consequences of colonial expansion. Yilan, located in northeastern Taiwan, is an example of a remote region where European control was limited and indirect for most of the time but was intensively involved in regional trade activities with foreign groups. In the mid-17th century, the VOC (Dutch East India Company) monopolized the deerskin trade in Taiwan for export on a large scale, involving many Indigenous communities and Han Chinese settlers [@Heyns2001; @Cheng2017]. Despite many discussions on the operations and policies of the deerskin trade through Dutch and Chinese written records [@Lin2014deer], direct investigations on the colonial experiences of local indigenous people through archaeological evidence are rare. The interaction between Indigenous people and their environments in the colonial setting of economic demands requires more Indigenous perspectives from archaeological studies.

<!-- research aim and questions -->
In this paper, I will investigate the Indigenous subsistence in a colonial context through the case study of the Kiwulan site (AD 1350-1950), one of the largest Iron Age settlements in northeastern Taiwan [@Chen2007]. Kiwulan encountered European colonial powers, including the Spanish and Dutch in the 17th century, followed by a large influx of Han Chinese migration in the 19th century, making it a suitable example to explore the reactions of Indigenous groups to the foreign presence over time. My research questions are: What was the resource use of Indigenous people before and after those foreign presences, especially during the Dutch period with large-scale trade? Can we identify the evidence of getting involved in the deer trade from faunal remains? If so, were there subsistence strategies in response to the impacts of economic demands by the Europeans and Chinese? To address these questions, I will analyze the faunal remains with a focus on vertebrate species from Kiwulan because zooarchaeology is one useful approach to studying local consumption and exchange systems [@Defrance2009; @Crabtree1990]. Through zooarchaeological evidence, this study aims to stress the active role of Indigenous people at the periphery of colonial bases in regional trade networks.

# The European colonization period and regional trade network in Northern Taiwan
<!-- Spanish colonization and the power relations in particular spatial and temporal contexts-->
In 1626, the Spanish (Spanish East Indies) established the first colony of San Salvador at Heping Dao, and later Santo Domingo in Tamsui in northern Taiwan (@fig-KWL-map), to counter the threat of the VOC (Vereenigde Oost-Indische Compagnie; Dutch East India Company) in the Asian market, which had established a colony in southern Taiwan two years earlier [@Borao2009, p.41; @Andrade2007]. Although the Spanish intended to build a colonial city, San Salvador remains its early development during their colonization, primarily consisting of a fortress and a church [@Borao2009, pp. 103-134; @Berrocal2018]. The neighboring area of the fortress was inhabited by two Indigenous groups, the Taparris and the Quimaurris, who had lived at the fort location before being expelled by the Spanish [@Borao2001, SIT:165, p.81]. The relations between the Spanish and Indigenous people might not involve extensive colonialism or intercultural domination, with limited territorial control only to the fort. Also, the Spanish relied on the exchange of local goods and food surpluses with Indigenous people, according to written records [@Borao2001, SIT:170-171]. This corresponds with a higher proportion of Indigenous and Chinese items with little European materials found in San Salvador [@Berrocal2018], indicating cultural entanglement.

<!-- the Dutch domination, how and why colonial powers were able to exert control -->
In 1642, the Dutch defeated the Spanish and took over San Salvador and Santo Domingo, expanding their colonial power to northern and eastern Taiwan until 1662. Compared to the Spanish, the Dutch was more strategic in consolidating their power by establishing political and economic institutions, forming alliances with Indigenous communities, and conquering resistant communities [@Kang2016; @Andrade2007]. For examples, the colonial institutions to Indigenous people include taxation, tribute system, village-leasehold system, census investigations, annual gatherings (Dutch: Landdag), and Indigenous representative from each aliened village [@Kang2016; @Andrade1997; @Andrade2007]. Among those institutions, the tribute system operated from 1642 to 1647 and the village-leasehold system introduced in 1644 had great impacts on Indigenous economy and the environment they lived in. Based on the tribute system, each household in allied communities was required to pay rice or deer products annually, especially deerskin. The amount of deerskin collected from Indigenous communities in northern Taiwan once reached 7,000 pieces in 1646 [@Andrade2007]. Moreover, to increase deerskin exports to Japan, the Dutch implemented a hunting license system (1636-1645) that gave Han Chinese permission to hunt deer on Indigenous lands during certain periods [@Heyns2001]. From 1633 to 1661, the average number of deerskins exported from Taiwan was about 67,000 per year [@Koo2011; @Cheng2017]. Those colonial activities might affect the economy of indigenous people since deer was an important resource in prehistoric Taiwan [@Chen2000animal].

<!--variation of colonial powers across regions and indirect colonization at Kavalan-->
Although the Dutch power expanded across nearly whole Taiwan island in the mid-17th century, the extent of their colonial dominance differed greatly from region to region. Colonial power over regions far from the European fortress were usually difficult and unsuccessful. Yilan in northeastern Taiwan is an example of minimal European dominance or indirect colonization for most of the time [@Kang2012]. As an plain located around 70 kilometers away from the European colonies with mountains in between, Yilan was hardly accessed by land but mostly from sea. The main Indigenous group in Yilan was the Kavalan. Given its strategic location on the route to Torrobouan, a place believed having gold mines, the Dutch was eager to integrate Yilan under their control [@DZI2000; @Kang2016, p47]. In 1644, the Dutch defeated several Kavalan communities and incorporated them into their colonial territory [@Kang2016]. Instead of engaging directly with the Kavalan people, the Dutch relied mostly on the assistance of the mediate, the Basay people active in northern Taiwan [@Kang2003], to coordinate and collect tributes from Yilan. This indirect approach allowed the maintenance of autonomy of the Kavalan people, and was even able to negotiate tributes requested by the VOC [@Kang2016].

<!-- the pre-existing trade network in norththern Taiwan-->
Before the arrival of the Europeans in northern Taiwan, an inter-insular trade network had been established since the 14th century, involving Chinese merchants and indigenous groups in Tamsui and Keelung [@Chen2005]. Chinese traders transported goods such as textiles, porcelains, beads, and metal items to exchange for raw materials from Taiwan, such as gold, sulfur, and animal hides and meats from the Basay people [@Chen2005; @Andrade2007]. This regional trade network was further connected to maritime trade between Ming China and the Ryukyu Kingdom during the 15th and 16th centuries and later to European long-distance trade in the 17th century [@Chen2005; @Von2019; @Wang2007]. Yilan was included in this long-established trade network, where the Kavalan people exchanged products such as rice and deerskins for foreign goods and handicrafts from the Basay people [@Kang2003]. This is evidenced by the large amount of imported foreign goods found across archaeological sites in Yilan, indicating frequent trade activities, especially around the 17th century [@TaiwanHistory2005]. This trade network was then intensified during the European presence, and the participation of Kavalan in this network was driven by the pursuit of foreign goods, and also likely colonial pressure from European powers. This unique historical and spatial context makes Yilan an ideal location to study Indigenous reactions to foreign influences and the associated trade networks from a long-term perspective in a pericolonial context [@Acabado2017; @Kang2016, pp. 31-78; @Trabert2017].

# Archaeological Context at Kiwulan
<!--introduction of Kiwulan and the population-->
Kiwulan, located in the northern part of Yilan (@fig-KWL-map), is a multi-component site, where the upper component spans from the late Iron Age to the historical periods (AD 1350-1850) [@Chen2007]. A large scale salvage excavation was conducted at Kiwulan that exposed a total of 262 excavation units (4 * 4 m), revealing abundant artifacts and numerous features, including middens, burials, ditches, and post-holes with wooden posts in situ. The excavtion suggests that Kiwulan was a well-developed settlement with longhouses (approximately 40 x 80 m) situated side by side and burials located to the north of the residential area [@Chen2004; @Chiu2004]. The main inhabitants of Kiwulan were the Kavalan people based on oral history and the similarities between Kavalan material culture recorded in historical records and archaeological artifacts [@Chen2004]. Kavalan people is the main Indigenous group in Yilan during the late Iron Age and the early modern period. According to a Dutch census conduced in the mid-17th century, there were about 40 Kavalan settlements with approximately 10,000 people in Yilan. Kiwulan, written as "Kibannoran" in Dutch records, was the largest settlement, with 160 households and 711 people [@Nakamura1938]. In the early 19th century, the population of Kiwulan dramatically decreased to fewer than 100 people due to the population pressure of Han Chinese migrants, which eventrually led to the relocation of Kavalan people to Hualien in eastern Taiwan [@Zhan1995].

<!--archaeological materials-->
The most common artifact found at Kiwulan was locally made pottery with geometric patterns as surface treatments, which was also found at many simultaneous sites in Yilan [@TaiwanHistory2005]. In addition to pottery, a variety of foreign goods, such as porcelains, stonewares, glass beads, carnelian beads, and metal artifacts and ornaments, were largely found in both the residential area and burials. Due to no evidence of metallurgy and local production of beads, these items are believed to have been acquired through exchange, primarily with the Basay people active in northern Taiwan. This indicates that Kiwulan has continuously participated in regional trade networks closely associated with northern Taiwan since the 14th century and influenced by the long-distance trade during the 17th century [@Wang2007]. Moreover, the use of foreign goods appears to have cultural significance, and previous studies suggested a possible increase in social differentiation due to the introduction of foreign goods [@Wang2020ornament; @Wang2021; @Cheng2008].

<!--archaeological remains-->
Abundant animal bones, shellfish, and plant remains were uncovered at Kiwulan, reflecting the diversity of natural resources in the plain and an environment close to estuary. The animal resource includes deer, suids, cattle, birds, fish, shellfish, and plants mainly consisting of rice, peach, pandanus, and cucurbits [@Chen2004]. Rice farming seems to be common but likely limited to a household scale rather than intensive agriculture according to historical records [@Chen1963]. Shellfish consist of bivalves such as clams and oysters and univalves from rivers and brackish water environments. Based on those archaeological remains and environments, Kiwulan residents should have practiced a mixed subsistence that includes rice farming, hunting, and fishing for wild resources. Artifacts associated with those subsistence practices were also found at Kiwulan, such as iron arrowheads for hunting, iron sickles for crop harvesting, and wooden paddles for water-related activities. However, the resource use across different time periods still need further investigations to better understands the subsistence adaptation relative to foreign presence and environmental changes.

```{r}
#| label: fig-KWL-map
#| echo: false
#| fig-cap: "The geographic location of Kiwulan and European fortress in northern Taiwan (adapted from Wang and Marwick, 2020). Map data from naturalearthdata.com."

knitr::include_graphics(here::here("analysis/figures/KWL_map.png"))
```

# Materials and Methods
The faunal assemblages from Kiwulan analyzed in this study were sampled from 40 adjacent units at the residential area excavated during the first season (2001-2003) [@Chen2007]. These 40 units were selected because of their location that were believed to be the major residential area with thick archaeological deposits and relatively intact archaeological contexts without disruption from modern construction. The spatial continuity can provides a consistent framework for understanding food practices over time on a settlement scale. Also, the same excavation and recovery methods applied at the site allow comparison between assemblages from different temporal sequences. I mainly focus on vertebrate species due to their economic values in the trade market. The contexts of faunal remains include dwelling spaces and middens that were associated with daily dietary practices, whereas those from burials that might relate to ritual activities are excluded.

To construct a reliable fine-grained chronological sequence, archaeological deposits were divided into six cultural layers (CL6-CL1) spanning from the 14th century to the 19th century based on excavation depth, stratigraphic sections, sediment color and texture, radiocarbon dates, and archaeological diagnostic materials that served as time indicators. The diagnostic materials include stoneware commonly found at European associated sites in early 17th-century Southeast Asia, such as Anping jars and martavan [@Sakai2012; @Ketel2011; @Dizon2016; @Hsieh1995; @Lee2021], pipes, and Chinese bricks relating to the 19th century [@Wang2011; @Hsieh2009]. Although precisely identifying the stratum representing the European arrival is challenging due to its short duration, the separation of strata into six layers based on the criteria mentioned above could still offer a relatively robust framework for comparing archaeological records before and after the foreign influence. Among these six temporal layers, the fourth cultural layer (CL4) can be associated with the 17th century, when the Europeans occupied northern Taiwan, while the second cultural layer (CL2) corresponds to the 19th century, associated with the large influx of Chinese migrants in Yilan.

Only the faunal assemblages identified to at least the family level were considered in this study for consistent comparison across time sequences. Deer antlers without attaching to cranium were excluded since they were less likely related to dietary practices. To obtain generalized pattern of animal use, measures of ubiquity and taxonomic abundance were applied. Ubiquity measures the frequency of specific taxa through their presence or absence in particular depositional contexts, suggesting the regularity of consumtion of that taxa [@Mckechnie2016; @Vanderwarker2010]. Here ubiquity values are expressed as the percentage of units with a given taxon to the total number of units per cultural layers. Faunal assemblages were primarily quantified using the Number of Identified Specimens (NISP) method, with the Minimum Number of Individuals (MNI) applied when necessary, such as for estimating the expected NISP at a site [@Grayson2004; @Lyman2012]. Different from MNI that determines the actual numbers of animals recovered from remains, NISP offers a broad overview of remains by considering all fragmented bone assemblages. Despite the criticism of overestimation the number of animals, NISP captures better variation in species abundance on a site scale with less impacts from sample size [@Grayson1984; @Lyman2008]. For comparison across animal classes, the normalized NISP were used to avoid potential biases in identification due to differences in the preservation or recovery of skeletal elements [@Grayson2004]. The normalized NISP is calculated by dividing NISP by the number of skeletal elements.

Given that deerskin was a primary commodity in local regional exchange and in high demand during the Dutch colonization period [@Cheng2017], the frequencies of skeletal parts and surface modifications of deer bones were examined to assess butchery and transport strategies. This can shed light on how indigenous people utilize deer in response to high economic demand when involved in regional trade network. Skeletal parts were quantified using NISP and classified into groups based on meat yields and portions: meaty (e.g., upperlimb), mid-meaty (e.g., vertebrae), and less meaty parts (e.g., skull and lower limb) [@Madrigal2004; @Munro2004]. Teeth without associated cranial elements were excluded to reduce overrepresentation due to their much higher preservation rate and quantities compared to other parts. While lower limbs and skull bones have minimal meat, they are often valued in many cultures for marrow and soft tissues, such as tongue and brain [@Madrigal2002]. Thus, less meaty elements were considered as additional source of nutrition in the interpretation. Similarly, the degree of fragmentation could also imply the extraction of marrow or grease [@Lupo1997], which is calculated by dividing the Minimal Number of Elements (MNE) by NISP.

Deer bone surface modifications relative to human butchering and cooking practices, such as cut marks, fracturing, and burning damage, were recorded, as well as natural marks such as weathering and root etching for understanding taphonomic effects. Cut marks on major limb joints, such as shoulder, elbow, wrist, hip, knee, and ankle [@Lyman2005], were quantified to explore butchering patterns over time. Moreover, cut marks on lower limbs relative to skinning processes were examined and compared. Those marks are characteristic of circular and straight incisions left by the series of cuts for acquiring intact animal hides [@Costamagno2019; @Isaakidou2007].

The data and R code used for these analyses and visualizations are openly available at xxx for research reproducibility and transparency (@Marwick2017; @MarwickandWang2022).

# Results

```{r setup, include= FALSE}
library(tidyverse)
library(knitr)
library(viridis)
library(cowplot)
```

```{r}
#| label: get-data
#| echo: false

# read data
kwl_fauna <- readxl::read_excel(here::here("analysis","data","raw_data","KWL_faunal_P.xlsx"), sheet = 1,
                                col_types = "text")

kwl_fauna_HM <- readxl::read_excel(here::here("analysis","data","raw_data","KWL_faunal_HM.xlsx"), sheet = 1,
                                   col_types = "text")

kwl_fauna_broken <- readxl::read_excel(here::here("analysis","data","raw_data","broken_fauna.xlsx"), sheet = 1,
                                   col_types = "text")

sample_p <- readr::read_csv(here::here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv"))

kwl_chro_6 <- readxl::read_excel(here::here("analysis", "data", "raw_data", "KWL_chronology.xlsx"), sheet = 5)

# tidy chronological data
kwl_chro_tidy_6 <-
  kwl_chro_6 %>%
  mutate(area = str_sub(Pit, 5, 7), Pit = str_sub(Pit, 1, 4)) %>%
  mutate(area = ifelse(area == "", "ABCD", area)) %>%
  mutate(area = sapply(strsplit(area, ""), paste, collapse= ",")) %>%
  separate_rows(area) %>%
  pivot_longer(cols = starts_with("L"), names_to = "layer", values_to = "period") %>%
  mutate(layer = str_sub(layer, 2, 3)) %>%
  mutate(layer = ifelse(nchar(layer) == 1, paste0("0", layer), layer))

# combine faunal data of cultural layers and temporal sequences
fauna_sam <-
  kwl_fauna %>%
  mutate(Pit = paste(context, pit, sep = "")) %>%
  filter(Pit %in% sample_p$the_sq) %>%
  filter(componant == "上文化層" &!layer == "00") %>% #remove sc
  #filter(!str_detect(`總報告類別`, "[[:punct:]]+|[0-9]+")) %>%
  mutate(area = ifelse(is.na(area), "C", area)) %>% # assign an area to the blanks, which won't change the results
  left_join(kwl_chro_tidy_6, by = c("Pit" = "Pit", "layer" = "layer", "area" = "area"))

# combine faunal data of middens and temporal sequences
fauna_H_sam <-
  kwl_fauna_HM %>%
  select(-`空間號`, -`層   位`) %>%
  mutate(Pit = paste("P", pit, sep = "")) %>%
  filter(Pit %in% sample_p$the_sq) %>%
  filter(componant == "上文化層") %>%
  left_join(kwl_chro_tidy_6, by = c("Pit" = "Pit", "layer" = "layer", "area" = "area"))

fauna_combined_context_all <-
  rbind(fauna_sam, fauna_H_sam) %>% # combine the two datasets
  filter(!`部位/名稱` %in% c("角","犄角","角?")) %>%  #not associated with diet
  filter(!`部位/名稱`== "角基部"|!is.na(`部位/左右`)) %>% # associated with cranial parts
  mutate(`重量(g)` = as.numeric(`重量(g)`)) %>%
  group_by(taxa) %>%
  mutate(`Weight (g)` = sum(`重量(g)`, na.rm = T)) %>%
  ungroup() %>%
  mutate(period = factor(period, levels = c("CL1","CL2","CL3","CL4","CL5","CL6"), order = T)) %>%
  mutate(category = case_when(str_detect(animal, "sika")|str_detect(animal, "sambar")|
                              str_detect(animal, "e deer")~ "deer", str_detect(animal, "pig")|
                                str_detect(animal, "boar") ~ "suid", str_detect(animal, "fish") ~ "fish",
                              str_detect(animal, "cattle")|str_detect(animal, "buffalo")~ "bovine",
                              animal == "aves" ~ "bird", TRUE ~ animal)) %>%
  mutate(class = case_when(str_detect(taxa, "muntjac")|str_detect(taxa, "ammal")|str_detect(taxa, "erv")|
                           str_detect(taxa, "Sus")|str_detect(taxa, "Rattus")|str_detect(taxa, "Rusa")|
                           str_detect(taxa, "Bos")|str_detect(taxa, "Bubalus")~ "mammal",
                           str_detect(category, "fish") ~ "fish", str_detect(category, "bird") ~ "bird",
                           TRUE ~ "reptile")) %>%
  mutate(`部位/左右` = case_when(`部位/左右` == "Ｌ" ~ "L", `部位/左右` == "Ｒ" ~ "R", TRUE ~ `部位/左右`)) %>%
  mutate(category = fct_relevel(category, "deer","suid","muntjac","bovine","rat","bird","fish","turtle")) %>%
  mutate(cutmarks = ifelse(str_detect(`人為痕跡`, "切"), "yes", "no"),
         fractures = ifelse(str_detect(`人為痕跡`, "削")|str_detect(`人為痕跡`, "砍")|
                            str_detect(`人為痕跡`, "折")|str_detect(`人為痕跡`, "敲"), "yes", "no"),
         burning =  ifelse(str_detect(`人為痕跡`, "火")|str_detect(`人為痕跡`, "燒"), "yes", "no")) %>%
  filter(is.na(refitted)) # for NISP

fauna_combined_context <- fauna_combined_context_all %>% filter(!taxa == "Rattus sp.")
```

```{r}
#| label: taxa-NISP-counts
#| echo: false
# taxa counts
fauna_combined_taxa <-
  fauna_combined_context_all %>%
  count(taxa, `Weight (g)`) %>%
  mutate(taxa = str_replace(taxa, "cervidae", "cervid")) %>%
  mutate(taxa = case_when(str_detect(taxa, "mm") ~ paste("Indeterminate", tolower(taxa), sep =" "),
                          TRUE ~ taxa)) %>%
  mutate(`Common name` = case_when(str_detect(taxa, "Cer") ~ "Sika deer", str_detect(taxa, "Bub") ~ 
                                   "Water buffalo", str_detect(taxa, "Rus") ~ "Sambar deer",
                                   str_detect(taxa, "Ree") ~ "Muntjac", taxa == "Sus scrofa" ~ "Suid",
                                   str_detect(taxa, "Bos") ~ "Yellow cattle", str_detect(taxa, "Rat") ~ 
                                   "Rat", str_detect(taxa, "Pli") ~ "Catfish", str_detect(taxa, "Act") ~ 
                                   "Ray-finned fish", str_detect(taxa, "Tes") ~ "Turtle",
                                   str_detect(taxa, "cervid") ~ "Deer", taxa == "Sus scrofa domestica" ~ 
                                   "Domestic pig", taxa == "Aves" ~ "Birds")) %>%
  mutate(group = case_when(str_detect(`Common name`, "deer") ~ 1, str_detect(`Common name`, "Deer") ~ 2,
                           str_detect(taxa, "Sus") ~ 3, str_detect(taxa, "Bub")|str_detect(taxa, "Bos") ~ 4,
                           str_detect(taxa, "munt") ~ 5, str_detect(taxa, "Rat") ~ 6, 
                           str_detect(`Common name`, "fish") ~ 7, str_detect(taxa, "Tes") ~ 8,
                           str_detect(taxa, "Ave") ~ 9, str_detect(taxa, "large m") ~ 10,
                           str_detect(taxa, "medium m") ~ 11, TRUE ~ 12)) %>%
  group_by(group) %>%
  arrange(desc(`Weight (g)`), .by_group = TRUE) %>%
  ungroup () %>%
  rename(Taxon = taxa, NISP = n) %>%
  select(Taxon, `Common name`,  NISP, `Weight (g)`) %>%
  bind_rows(summarise(., across(where(is.numeric), sum), across(where(is.character), ~"Total"))) %>% 
  mutate(`Common name`= ifelse(is.na(`Common name`)|`Common name` == "Total", "-", `Common name`)) %>% 
  mutate(Taxon = ifelse(Taxon == "Sus scrofa", "Sus spp.", Taxon))

total_NISP <- fauna_combined_taxa %>% select(NISP) %>% tail(1) #rev(fauna_combined_taxa$NISP)[1]
unident <- fauna_combined_taxa %>% filter(str_detect(Taxon, "In")) %>% summarise(sum(NISP))
deer_NISP <- fauna_combined_taxa %>% 
  filter(str_detect(`Common name`, "eer")) %>% summarise(sum(NISP)) #|str_detect(Taxon, "Ree")
suid_NISP <- fauna_combined_taxa %>% filter(str_detect(Taxon, "Sus spp.")) %>% summarise(sum(NISP))
buff_NISP <- fauna_combined_taxa %>% filter(str_detect(Taxon, "Bub")) %>% summarise(sum(NISP))
deer_NISP_per <- round(deer_NISP/(total_NISP-unident), 2)*100
suid_NISP_per <- round(suid_NISP/(total_NISP-unident), 2)*100
buff_NISP_per <- round(buff_NISP/(total_NISP-unident), 2)*100
```

The faunal assemblage analyzed in this study from Kiwulan is composed of `r total_NISP` bone fragments based on NISP (@tbl-counts). Among the identifiable bones, large deer is the majority (`r deer_NISP_per`%), followed by suid (`r suid_NISP_per`) and water buffalo (`r buff_NISP_per`). To further compare general patterns of animal consumption in terms of their economic values over time, animals in the same family with similar body size were grouped together. Thus, the deer category includes sika deer, sambar deer, and large deer, while water buffalo and yellow cattle were group into the category, bovine. Similarly, domestic pig (n = 2, from CL4 and CL6) was integrated into the suid category. Rats (Rattus sp.) were excluded from further analyses as they are normally considered commensal animals, and their relatively intact skeletons from one single archaeological context show no direct evidence of being part of the daily diet. In general, the faunal remains show the frequencies increased from CL6, peaked at CL4, and then declined dramatically after CL4 ( (@tbl-counts, @fig-NISP-proportion: A). A similar trend was also observed in the weight of unidentified small bone fragments (@fig-NISP-proportion: B).

```{r}
#| label: tbl-counts
#| echo: false
#| tbl-cap: "Taxa, Number of Identified Specimens (NISP), and weight of faunal remains from the sampled area at Kiwulan."
#| tbl-colwidths: [40,25,15,20]
#| tbl-cap-location: bottom
library(kableExtra)
knitr::kable(fauna_combined_taxa, align = "l") 
```

```{r}
#| label: mammal-counts
#| echo: false
mammal_count <-
  fauna_combined_context %>%
  filter(class == "mammal") %>%
  count(period, category) %>% 
  drop_na() %>% 
  bind_rows(summarise(., across(where(is.numeric), sum), .by = period)) %>% 
  mutate(category = as.character(category)) %>% 
  mutate(category = ifelse(is.na(category), "total", category))

CL4 <- mammal_count %>% filter(period == "CL4") 
CL4_boar <- round(CL4$n[3]/CL4$n[4], 3)*100

# taxonomic diversity
category_count <-
  fauna_combined_context %>%
  count(period, category) %>% 
  count(period, category) %>% 
  drop_na() %>% 
  bind_rows(summarise(., across(where(is.numeric), sum), .by = period)) %>% 
  slice(24:29) %>% 
  select(period, n) %>% 
  rename(diversity = n)

# wide format and join
wider_mammal_count <-
  mammal_count %>% 
  pivot_wider(id_cols = period, names_from = category, values_from = n, values_fill = 0) %>% 
  left_join(category_count)

boar_reg <- summary(lm(suid ~ total, data = wider_mammal_count))
boar_lm <- c(round(boar_reg$adj.r.squared, 2), sprintf('%.3f', boar_reg$coefficients[,4][2]))
munt_reg <- summary(lm(muntjac ~ total, data = wider_mammal_count))
munt_lm <- c(round(munt_reg$adj.r.squared, 2), sprintf('%.2f', munt_reg$coefficients[,4][2]))
deer_reg <- summary(lm(deer ~ total, data = wider_mammal_count))
div_reg <- summary(lm(diversity ~ total, data = wider_mammal_count))
div_lm <- c(round(div_reg$adj.r.squared, 2), sprintf('%.2f', div_reg$coefficients[,4][2]))
```

For animal classes, the relative abundance shows that mammals are the primary dietary source throughout all layers, with fish as the secondary source in CL4 and birds in CL3 and CL2 (@fig-NISP-proportion: C). Fish and birds are both identified from CL5 to CL3; however, their proportions (less than 10%) are much lower compared to mammals. The reptile is represented by one piece of turtle shell found in CL4, which might not be a common food content at the site. Within those identifiable mammals, deer is the main animal from CL6 to CL2 that accounts for at least 85% per layer, while CL1 is predominated only by bovine (@fig-NISP-proportion: D). Suid was present from CL6 to CL3, and had a higher proportion in CL4 (`r CL4_boar` %), while muntjac appeared in CL5, and was continuously present until CL2. In general, there is an increase in animal diversity in CL4 and CL5, especially in CL4, where suid and fish have slightly higher proportions compared to those in other layers. Given that the taxonomic richness, or animal diversity, could be impacted by sample size varied from layer to layer, the relationship between the number of taxa and sample size (NISP) was examined by linear regression model. The result shows no correlation between the diversity of animals and sample size ($R^2$ = `r div_lm[1]`, P = `r div_lm[2]`), which suggests little bias from potential sample size.

```{r}
#| label: fig-NISP-proportion
#| echo: false
#| fig-cap: "A: The number of identified specimen per cultural layers. The NISP data is derived from Table 1 . CL6 represents the earliest layer and CL1 is the youngest. B: The weight of the unidentified small bone fragments for each cultural layer. C: Relative abundance of vertebrate classes by cultural layers based on normed NISP. D: Relative abundance of identified mammals by cultural layers based on NISP."

knitr::include_graphics(here::here("analysis/figures/NISP_pro_by_layer.png"))
```

```{r}
#| label: tbl-ubiquity
#| echo: false
#| tbl-cap: "Ubiquity of faunal categories across the sampling area at Kiwulan."
#| tbl-cap-location: bottom
# calculate the number of units for each cultural layers
kwl_unit_per_period <-
  kwl_chro_6 %>%
  mutate(all = pmap_chr(select(., -Pit, -Grid), ~toString(unique(na.omit(c(...)))))) %>%  # unite and remove duplicates
  mutate(unit = str_extract(Pit, "[A-Z]+[0-9]{3}")) %>%
  distinct(unit, .keep_all = TRUE) %>%
  mutate(CL1 = ifelse(str_detect(all, "CL1"), unit, NA)) %>%
  mutate(CL2 = ifelse(str_detect(all, "CL2"), unit, NA)) %>%
  mutate(CL3 = ifelse(str_detect(all, "CL3"), unit, NA)) %>%
  mutate(CL4 = ifelse(str_detect(all, "CL4"), unit, NA)) %>%
  mutate(CL5 = ifelse(str_detect(all, "CL5"), unit, NA)) %>%
  mutate(CL6 = ifelse(str_detect(all, "CL6"), unit, NA))

# count the units by cultural layers for later join
unit_by_period <-
  data.frame(unit_count = colSums(!is.na(kwl_unit_per_period))) %>%
  slice_tail(n = 6) %>%
  tibble::rownames_to_column("period")

# ubiquity of taxa from the 40 units by temporal sequences
fauna_ubiquity <-
  fauna_combined_context %>%
  select(Pit, layer, category, `重量(g)`, period, taxa) %>%
  count(period, Pit, category) %>%
  count(period, category) %>%
  drop_na(period, category) %>%
  left_join(unit_by_period) %>% # join the unit counts for each period
  mutate(ubiquity = paste(round(n/unit_count *100, 2), "%")) %>%
  select(-n, -unit_count) %>%
  pivot_wider(names_from = period, values_from = ubiquity) %>%
  mutate(across(CL1:CL6, ~replace_na(.x, "-"))) %>% 
  arrange(category)

CL4_ubi_deer <- fauna_ubiquity$CL4[1]
CL4_ubi_boar <- fauna_ubiquity$CL4[2]
CL4_ubi_munt <- fauna_ubiquity$CL4[3]

# get the bone mineral desity data
deer_element_BMD <- readr::read_csv(here::here("analysis", "data", "derived_data", "deer_BMD.csv"))
cor_sur_BMD <- cor.test(deer_element_BMD$NISP, deer_element_BMD$BMD)
survivor_cor <- round(c(cor_sur_BMD$estimate, cor_sur_BMD$p.value),2)

# for deer
fauna_deer_only <-
  fauna_combined_context %>%
  filter(category %in% c("deer") & !is.na(period))

deer_NISP <- fauna_deer_only %>% count(period) %>% rename(NISP = n)

# calculate MNE
deer_MNE <-
  fauna_deer_only  %>%
  filter(!is.na(`部位/左右`)&is.na(overlap)) %>% # remove overlapped fragments
  count(period,`部位/左右`,`部位/名稱`,`部位/位置`) %>%
  group_by(period) %>%
  summarise(MNE = sum(n), across())

deer_NISP <- fauna_deer_only %>% count(period) %>% rename(NISP = n)
frag_index <- left_join(deer_MNE, deer_NISP) %>% mutate(index = MNE/NISP) %>% distinct(period, index) %>% 
  mutate(index = round(index, 2))
frag_CL3_CL4 <- c(frag_index$index[2], frag_index$index[3])

knitr::kable(fauna_ubiquity)
```

@tbl-ubiquity shows deer had been the most widespread animal from CL6 to CL2, with the highest ubiquity in CL4 and CL5 (`r CL4_ubi_deer`), and lowest in CL2. The second and third animal regularly used are suid and muntjac. Although those two animals reached their highest ubiquity in CL4 (`r CL4_ubi_boar` for suid, and `r CL4_ubi_munt` for muntjac), their presence patterns are slightly different. Suid was utilized as early as deer in CL6, declined in CL3, and disappeared in CL2, whereas muntjac showed a cultural lag, appearing one phase later than suid. Bovine had a low ubiquity in CL5 and high ubiquity in CL6. Small animals (bird, fish, and turtle) had much lower ubiquity than large animals and they had a slightly increase in ubiquity in CL3. The result of animal ubiquity correspond with their frequencies (NISP), suggesting that deer is the most common resource at the site, with an increase utilization of suid and muntjac in CL4.

Although deer was predominant throughout all temporal sequences, variations in their skeletal parts could be identified. Deer bones were grouped into six body portions [@Madrigal2004], including head (cranium, mandible), axial (vertebrae, sternum, ribs), upper forelimb (scapula, humerus, ulna, and radius), upper hindlimb (femur, tibia, innominate), lower limb (carpals, metacarpals, tarsals, metatarsals, astragalus, calcaneus), and foot (phalanges). The result (@fig-deer-portion: A) shows that the meaty portion, especially the upper forelimb, is the most frequent part from CL6 to CL2, followed by upper hindlimb and the less-meaty portions, lower limb and foot. The consistency of higher numbers of upper limb and lower limb suggests that those parts should be transported together [e.g., @Stiner1991]. The generally low frequency of axial elements might be likely due to a high degree of fragmentation, butchery practices, or dietary preferences. A different pattern between layers is observed in the frequency of cranial bones, for which a few present in CL6, increased to the highest number in CL4, and then decreased thereafter. In addition, lower limb and foot bones appear to be proportionally less in CL2 and CL3 compared to earlier periods. Those differences might indicate changes in processing decisions of meat portions of deer.

Considering the impacts of differential bone density on preservation of skeletal elements, the relationship between the survivorship of deer skeletal elements and bone mineral density was tested by Pearson correlation. The values of bone survivorship were collected from @Lyman1994's research on deer, and calculated by dividing NISP by expected NISP, followed by multiplying 100%. The result (@fig-deer-portion: B) shows no significant positive correlation ($r$ = `r survivor_cor[1]`, P = `r survivor_cor[2]`)) between bone survivorship and bone density, suggesting that density-mediated bias has little impact on the taphonomic processes of bones. For the degree of bone fragmentation, CL3 shows higher fragmentation (index = `r frag_CL3_CL4[1]`), while bones in CL4 are less fragmented (index = `r frag_CL3_CL4[2]`). However, the fragmentation rates are not significantly differed between layers.

```{r}
#| label: fig-deer-portion
#| echo: false
#| fig-cap: "A: The frequency of skeletal parts of deer using NISP across cultural layers at Kiwulan. B: Bone minerial density agaist %survivorship of deer remains."

knitr::include_graphics(here::here("analysis/figures/deer_elements.png"))
```

```{r}
#| label: tbl-deer-modify
#| echo: false
#| tbl-cap: "Number and proportion of deer bones with surface modifications at Kiwulan."
#| tbl-cap-location: bottom
#| tbl-colwidths: [10,20,20,20,20,10]

# bones with surface modifications
burned_deer <- fauna_deer_only %>% filter(burning == "yes") %>% count(period) %>% rename(burned = n)
cutmark_deer <- fauna_deer_only %>% filter(cutmarks == "yes") %>% count(period) %>% rename(cutmarks = n)
frac_deer <- fauna_deer_only %>% filter(fractures  == "yes") %>%
  filter(`部位/名稱` %in% c("脛骨","股骨","肱骨","橈骨","上肢骨","掌骨","掌骨或蹠骨","蹠骨")) %>%
  count(period) %>% rename(fractures = n)
post_deposit <-fauna_deer_only %>% filter(!is.na(`保存狀態`)) %>% count(period) %>%
  rename(`post-deposition` = n)

list_modif <- list(burned_deer, cutmark_deer, frac_deer, post_deposit, deer_NISP)

# join
deer_modify <-
  list_modif %>%
  reduce(inner_join, by = 'period') %>%
  mutate(cut_per = round(cutmarks/NISP, 3) * 100, burn_per = round(burned/NISP, 3) * 100,
         frac_per = round(fractures/NISP, 3) * 100, post_per= round(`post-deposition`/NISP, 3) * 100) %>%
  mutate(burned = paste(burned, paste0("(", burn_per, " %)")),
         cutmarks = paste(cutmarks, paste0("(", cut_per, " %)")),
         fractures = paste(fractures, paste0("(", frac_per, " %)")),
         `post-deposition` = paste(`post-deposition`, paste0("(", post_per, " %)"))) %>%
  rename(`cut marks` = cutmarks) %>%
  select(-cut_per, -frac_per, -burn_per, -post_per)

knitr::kable(deer_modify)
```

Surface modifications, including cut marks, burning damage, and fractures were identified on deer bones from CL6 to CL2, with variations in their proportions (@tbl-deer-modify). Cut marks were present on 50% of bones in CL2, 30% of bones in earlier period (CL6 to CL4), and at a lower proportion of 23% in CL3. Most cut marks were located on major limb joints and long bones, suggesting butchering activities related to disarticulation and meat detaching. @fig-deer-cutmark shows that at least 50 % of joints in CL4 present cut marks, with a higher proportion of 75% at the hip. A similar pattern was observed in CL5, but with a lower proportion of cut marks on the ankle, knee, and hip compared to CL4. Not all joints were found in CL6, CL3 and CL2, making direct comparisons across all layers difficult. Nevertheless, the overall high proportion of joints with cut marks suggests intensive processing of deer throughout the site's occupation. In addition, cut marks on lower limbs should be associated with skinning (@fig-cutmark-img: e) and were primarily found on metapodials in CL6 (n = 1), CL5 (n = 4), CL4 (n = 6), CL3 (n = 1), followed by first phalanges in CL4 (n = 1) and CL5 (n = 1). Some joints displayed clean, smooth cut surfaces from chopping (@fig-cutmark-img: a, c), along with narrow V-shaped cut marks observed on most bones, indicating the use of metal tools [@Greenfield1999].

Burned bones (@fig-cutmark-img: d), characterized by charring and color changes, had the highest proportion in CL6 at 40% and gradually decreased to less than 10% after CL4. Bone breakage marks, including spiral fractures, transverse fractures, longitudinal splits, and percurssion scars [@Haynes1983; @Gifford2018; @Fisher1995], were observed frequently on long bones across different layers (@fig-fracture-img). These fractures indicate intentional marrow extraction at Kiwulan. CL2 had the highest proportion of fractured bones (> 50%), compared to 35% in CL3 and CL6, and 25% in CL4 and CL5. Chewing marks were observed on a few bones represented by a series of superficial elongated scores and circular depressions (@fig-fracture-img: e)[@Bello2023], suggesting carnivore gnawing. In addition, natural alterations to bones resulting from post-deposition impacts were also observed, such as weathering and root etching [@Gifford2018natural], on approximately one-third of the bones across all layers (@fig-fracture-img: c,d; @tbl-deer-modify). Those post-depositional impacts might have led to an underestimation of cut marks in general.

```{r}
#| label: fig-deer-cutmark
#| echo: false
#| fig-cap: "The proportion of cut marks on major limb joints of deer across cultural layers. The method for clasifying and quantifiying limb joints is based on Lyman (2005)."

knitr::include_graphics(here::here("analysis/figures/deer_cutmarks.png"))
```

```{r}
#| label: fig-cutmark-img
#| echo: false
#| fig-cap: "Examples of deer bones with cut marks, chop surfaces, and burning damage. Cut marks are indicated by red squares, and chop surfaces are indicated by black arrows. Bone fragments are (a) proximal radius with cut marks and a chop surface, (b) distal humerus with cut marks, (c) acetabulum with chop surfaces, (d) charred astragulus, (e) metatarsal with cut marks."

knitr::include_graphics(here::here("analysis/figures/deer_cutmark_img.png"))
```

```{r}
#| label: fig-fracture-img
#| echo: false
#| fig-cap: "Examples of deer bones with fractures, natural modifications, and chewing marks. Bone fragments are (a) proximal metatarsal with a puncturing hole and longitudinal split, (b) distal tibia showing spiral fracture, (c) distal radius showing tranverse fracture and root etching, (d) femur shaft showing longitudinal split, weathering, and a percurssion scar indicated by an arrow, (e) scapula with chewing marks."

knitr::include_graphics(here::here("analysis/figures/deer_fracture_img.png"))
```

# Discussion
<!--results and chronology -->
The sampled faunal assemblage from the major residential area at Kiwulan suggests a continuation of sika deer consumption from CL6 to CL2, and then shifted to bovine in CL1. Despite deer was a substantial part of subsistence at Kiwulan, variations in their quantity, ubiquity, butchering patterns, and proportions relative to other animals show changes that might relative to foreign impacts. Based on radiocarbon dates and diagnostic artifacts indicating specific time [@Wang2020ornament; @Chen2007], the cultural layers were further grouped into temporal sequences in relation to foreign presences for further discussion. The temporal sequences include the pre-European period (CL6, CL5), the European period (CL4), the post-European period (CL3), and the arrival of Han-Chinese immigrants (CL2, CL1). It should be noted that the European period does not mean a solid European colonial control over Kiwulan, but a broad impact from European occupation in Northern Taiwan and associated colonial economic activities on indigenous people in Yilan, such as getting involved in the long-distance trade network, tribute paying, and suffered from occasional European attacks. The small sample size after separating them into different periods might limits a comprehensive interpretation. However, comparing the results in a continuous sequences still provide evident patterns reflecting changes in animal use.

<!--summary of overall observations across time periods -->
The sampled area at Kiwulan shows a large quantity of faunal remains in the European period, followed by the late pre-European period. Interestingly, deer bones from the late pre-European period presented the same ubiquity as in the European period but in only half the quantity. This might indicate consistency in the  regularity of deer consumption at the settlement scale before and after European presence, but deer utilization was intensified in the European period. During the European period, in addition to the predominance of deer, an increased relative proportion and ubiquity of suid suggests a more regular use of suid than in other periods. The dramatic decrease in deer bones after the European period might be related to both the decline of the Kiwulan population and a decrease in the wild deer population. Sika deer populations in western Taiwan have significantly shrunk since the 18th century because of habitat loss and overhunting [@Chen2000animal]. An increase in muntjac use may suggest they served as a substitute, replacing the role of suids in earlier periods. Despite this, deer remained a predominant resource for local consumption at Kiwulan during the post-European and early Chinese period. Bovine appeared in the pre-European period with just a few evidence (yellow cattle), but became the only animal (water buffalo) used at Kiwulan in the late Chinese period. Since bovine was not domestic aniamls that were beleived introduced from mainland China, this suggests a Chinese influence and the relocation of the Kavalan people to Hualien due to population pressures from Chinese immigration [@Zhan1995].

<!--skeletal part frequencies and butchery patterns of deer-->
Although deer was the primary resource at Kiwulan, differences in skeletal completeness across periods suggests changes in butchery locations and cooking practices. There were more complete carcasses in the European and late pre-European period compared to the post-European and Chinese periods. Moreover, the substantial rise in fragmented head parts (cranium and mandible) during the European period indicates greater use of soft tissues, such as brains and tongues, or the transport of more intact deer to the settlement for processing. This aligns with the historical account in 1603, Record of Formosa (Dongfan ji) [@Chen1603], which describes deer tongue consumption and group hunting for deer in western Taiwan. Changes in cooking practices are also evident. The decline in burned bones over time with a lowest percentage in the European period suggests a shift from roasting to other cooking methods, such as boiling. The presence of small chopped bones that could fit in local ceramics at Kiwulan indicates boiling as a possible practice since the European period. This could be further confirmed through organic residue analysis on pottery in the future. Other meat preparation such as drying (or salting and smoking) cannot be ruled out [e.g., @Spyrou2019]. Although deer vertebrae and ribs, portions potentially used for drying [@Friesen2001; @Binford1978], were in low frequencies at Kiwulan, dry meat consumption or exchange by the Kavalan people was reported in historical records [@Chen1963]. The shift in cooking practices might result from increased need for deer, which could have led people to adopt boiling due to its additional benefits, such as maximizing residual meat use, facilitating grease rendering, and improving quality of lean meat [@Lupo1997], and drying for meat preserving and transporting.

A consistent pattern in deer part frequencies was also identified throughout all periods. There were high frequencies of upper limb bones but low frequencies of vertebrae and ribs, suggesting selectively butchered and preferences of meaty portions. Moreover, among those meaty portions, front quarters (humerus and radius) are always more abundant than hindquarters (femur and tibia). One possible explanation is that femurs and tibiae were heavily fragmented for marrow extraction due to their high yields [@Madrigal2002; @Lupo1998], making them less identifiable in the assemblage. Marrow extraction can be further supported by at least 25% of long bones showing fractures and a moderate degree of fragmentation in general. Another likelihood is that some hindquarters might be used for trading, as Dutch records mention deer legs and meats as commodities in addition to hides [@Blusse2006, pp.220-224], but without specifying which legs. Deer legs imply bone-meat packages that would result in a low frequencies of bones of hindquarters or front quarters. Considering the overall high proportion of meaty portions at Kiwulan, deer were likely processed not only for local consumption but also for trade. The lower frequencies of hindquarters than front quarters might be attributed to both marrow extraction and export of deer products, rather than a sole contribution.

<!--deer hides-->
Deerskin trade had prevailed since the 14th century in East Asia, and monopolized by the Europeans in the 17th century, when large hides were exported to Japan, especially from western Taiwan [@Cheng2017; @Hollmann1991; @Wang1981]. The cut marks identified on deer lower limbs at Kiwulan confirm the removal of deer skins and likely being traded with external traders. Abundant foreign goods at Kiwulan, including ornaments, porcelains, and metal items, supports frequent exchanges between local residents and outsiders [@Wang2020ornament; @Chen2007]. This suggests that not only western Taiwan but also northeastern Taiwan was intensively involved and impacted by the long-distance trade. Although lower limbs with skinning cuts did not show a substantial increase in the European period, the disproportionate increase in head parts might hint at a likelihood of brain tanning when preparing hides in addition to tissue consumption. Brain tanning is a traditional way for softening and preserving hides in many Indigenous cultures, such as North America, Southern Africa, North Asia, and Japan [@Baillargeon2010; @Badenhorst2009; @Hatt1969; @Matsumoto2016]. The present-day Bunun group in the central highlands of Taiwan has revived hide tanning based on oral history, primarily using peanuts or pig lard as tannnig materials, with a few examples using animal brains [@Li1999, pp. 48; @Xu1994]. Despite the materials for tanning used by the Kavalan people being unknown due to the loss of traditions, brain tanning is a practicable method given the availability of deer heads from hunting.

<!-- population dynamics and arguments-->
The high population during the European period inferred by pottery could contribute to the increasing animal use overall [@Wang2020pot; @Hsieh2009]. However, the impacts from foreign presence can be still identified through deer butchery patterns, relative proportions of animals, and taxon ubiquity. First, lower limbs with skinning cuts were mostly in the European period, followed by the late pre-European phase, corresponding with the rise of regional trade in the 14-15th century and was later intensified in the 17th century. Second, if the high frequencies of deer parts, especially heads for brains or tongues consumption, during the European period solely resulted from food stress due to population growth, a parallel rise in marrow extraction would be expected. However, long bones with fracture patterns were 10-30 % lower during the European period and the previous phase than in other periods. Third, population growth should have led to a more proportional increase in both large and small animals. Despite a high taxonomic diversity during the European period, suid presents a disproportionate rise in their ubiquity, quantity, and proportion. This indicates that deer products trade may have driven Indigenous people to rely more on suids as an alternative prey, likely because of their comparable body size and meat return rates to deer [@Broughton2011prey]. Suids as a secondary resource after deer in Indigenous cultures were observed in western Taiwan [@Chuang2021]. This hints the role of suid as a supplement when deer was not sufficient or available. After the Dutch left, deerskin trade continued but in a small scale with around 10,000 pieces exported annually, which is much less than 100,000 pieces in the European period [@Nakamura1997; @Tsao2011]. The small trade market and low population at Kiwulan might lead to a reduced reliance of suid, as shown by a low proportion of suid in the post-European period and eventually disappeared during the Chinese period.

<!--indigenous reaction and resilience-->
Compelling the zooarchaeological evidence and descriptions from historical records, the substantial increase in deer bones in the European period not just simply the result of population growth, but reflect intensified hunting driven by deer trade. Deer is the first priority of food source and commodities at Kiwulan since the late pre-European period and their economic benefits were intensified during the European period. It should be recognized that the role of the Dutch is both colonist who demanded tributes from Yilan, and another trader who were incorporated into the pre-existing inter-insular trade involving Han Chinese merchants and other indigenous groups in Northern Taiwan [e.g., @Torrence2000]. Local Indigenous people might take a more active participation in the trade rather than resisted due to the benefits of obtaining foreign necessities (e.g. metal knives and pan) and goods (e.g. ornaments and porcelains) that were highly valued in local cultures [@Wang2020ornament]. Despite an indirect and weak Dutch colonial control in Yilan, the stress from colonial powers cannot be totally neglected given the local experiences of the Dutch attacks occasionally. Everyday practices play an essential role for maintaining Indigenous traditions within a colonial setting or frequent cultural exchange [@Silliman2009; @Noe2023]. The persistence of deer consumption with similar butchery patterns over time at Kiwulan reflect a maintained tradition even after other animals were introduced, for example, bovine in the pre-European period. The increased reliance on suid during the European period could be viewed as a strategy to accommodate the high demands of deer products from trade network.

# Conclusion
By contextualizing zooarchaeological evidence into time periods relative to foreign presence, including the Europeans in the 17th century, and Chinese immigrants from the late 19th century, this study reveals the participation of Indigenous groups in regional trade work and adaptation strategies. The frequencies, butchering patterns, and human modifications on deer bones confirms the active deerskin trade since the 14th century and intensified during the European period. The results demonstrates even Indigenous people in the peripheral area with relatively weak colonial control were also profoundly impacted by colonial activities than previously thought. The continued prioritization of deer consumption at Kiwulan suggests a cultural persistence of food practices, while the strategic reliance on suids during the European period demonstrates adaptation to high demands for deer products, and later incorporating more muntjac in response to overhunting of sika deer. This highlights how Kiwulan residents reacted to foreign presence and environmental changes and demonstrated cultural resilience. The long-term dynamics of animal use in northeastern Taiwan can enhance our understanding of Indigenous reaction within a specific social-cultural context. The role of Indigenous people might be more active in the exchange network considering the large amount of foreign goods that were incorporated into local cultures and used as prestige items. A more comprehensive investigation across Indigenous settlements in northeastern Taiwan and including other dietary evidence, such as botanical remains and organic residue analysis in the future could provide a deeper insights into the regional dietary patterns and human-environment dynamics.

# Acknowledgements

I would like to thank the Yilan County Cultural Affairs Bureau in Taiwan for permitting access to the zooarchaeological remains used in this study and the archaeological team for the early documentation of faunal remains that provides a foundation for this study. Special thanks to Chien-Hsiang Lin at the Biodiversity Research Center, Academia Sinica, for his guidance and assistance in identifying fish bones. I also appreciate Jui-Yang Tsou for his valuable comments and assistance during the identification process. I would like to express my gratitude to Yin-Ru Chiang and his team at the Biodiversity Research Center, Academia Sinica, for providing me a lab space to work on data and finilizing this work. I am also grateful to the reviewers of this manuscript for their comments.

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
#| label: colophon
#| cache: false

# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```