### 2025-03-09 analysis/paper/paper.qmd
draft editing

---
title: "Indigenous Subsistence and Colonial Encounters: A Zooarchaeological Perspective from Kiwulan in Northeastern Taiwan"
author:
  - Li-Ying Wang:
      correspondence: "yes"
      email: liyingw15@gmail.com
      orcid: 0000-0001-9060-2125
      institute: Academia Sinica
title-block-published: "Last updated"
date: now
date-format: long
format: 
  docx:
    reference-doc: "../templates/template.docx" # Insert path for the DOCX file
execute:
  echo: true
  warning: false
  message: false
  comment: "#>"
  fig-path: "../figures/"
  fig-dpi: 600
filters:
  - ../templates/scholarly-metadata.lua
  - ../templates/author-info-blocks.lua
  - ../templates/pagebreak.lua
bibliography: references.bib
csl: "../templates/journal-of-archaeological-science.csl" # Insert path for the bib-style
abstract: |
  Colonial encounters between Europeans and Indigenous peoples often had profound impacts on Indigenous societies and cultures in many parts of the world. In northeastern Taiwan during the Late Iron Age, local Indigenous groups experienced intense foriegn contact through regional trade networks in East Asia, which expanded further with the arrival of Europeans in the 17th century. This study focuses on local subsistence through faunal remains to explore how Indigenous communities responded to trade activities driven by Europeans and Han Chinese, particularly the high demand for natural resources such as deer. By analyzing faunal remains from the upper component of Kiwulan (AD 1350-1850), a late Iron Age site in northeastern Taiwan, this research dicusses the Indigenous participation in the deerskin trade and their adaptive strategies. The results suggest cultural persistence in deer consumption, along with increased reliance on suids during the European period as demand for deer products intensified. This paper also provides insights into colonial experiences from Indigenous settlements at the periphery of European colonial territories, a perspective rarely explored.

keywords: |
  Zooarchaeology; Iron Age; Taiwan; sika deer; trade; European colonization
highlights: |
  Indigenous community in northeastern Taiwan engaged in regional deer trade were identified from faunal remains.
  The high freqiencies of deer bones at Kiwulan reveals continously cultural persistence in deer consumption.
  Changes in deer butchery patterns and cooking practices indicates an intensified hunting during the European colonization period.
  Increased reliance on suids during the European period suggests adaptive strategies to sustain local subsistence.
---

<!-- This is the format for text comments that will be ignored during renderings. Do not put R code in these comments because it will not be ignored. -->

<!-- With the following code you can access and display values from the yml header above. -->

Keywords: `r rmarkdown::metadata$keywords`

Highlights: `r rmarkdown::metadata$highlights`

<!-- The actual document text starts here: -->

# Introduction
<!-- introduction about European expansion -->
European expansion since the 15th century was closely tied to long-distance trade, exploration, and colonization, which often had profound impacts on Indigenous societies. The impacts also varied across regions depending on specific geographic and historical circumstances, resulting in different forms of European colonization and power relations [@Jordan2009]. Understanding the diverse and dynamic colonial processes has become a key focus in archaeology, because it provides critical insights into the complex consequences of European expansion [@Gosden2004]. In East Asia, European colonization was mostly driven by the pursuit of raw materials, natural resources, or luxury goods for commercial purposes [@Prakash1997]. The new exchange systems and colonial institutions introduced by Europeans profoundly influenced local economies, social structures, or settlement patterns [@Christie1996; @Junker2004; @Lape2002]. However, recent studies emphasize that Indigenous groups were not just passively accept European colonization; instead, they actively responded through resistance, negotiation, and cultural persistence or resilience, leading to complex and entangled colonial experiences [@Torrence2003; @Silliman2005; @Mitchell2010; @Jordan2009; @Alexander1998; @Noe2023].

<!-- NE Taiwan as a case study -->
While many studies focus on major colonial centers, peripheral regions also experienced varying degrees of European influence through colonial powers and associated trade networks, namely pericolonialism [@Acabado2017; @Trabert2017]. Exploring these areas is important to for understanding the dirverse impacts of colonial expansion. Yilan, in northeastern Taiwan, is an example of a remote region where European control was limited and indirect but Indigenous communities were deeply involved in regional trade with foreign groups. In the mid-17th century, the VOC (Vereenigde Oost-Indische Compagnie; Dutch East India Company) monopolized the deerskin trade in Taiwan and exported deerskin on a large scale to Japan [@Heyns2001; @Cheng2017]. While historical records by the Dutch and Chinese provide details about the operations and policies of the deerskin trade [@Lin2014deer], archaeological investigations into Indigenous reactions to these economic activities remain rare. To fully understand Indigenous responses to colonial economic demands, archaeological perspectives on their interactions with the environment are essential.

<!-- research aim and questions -->
In this paper, I investigate Indigenous subsistence in a pericolonial context through the case study of the Kiwulan site (AD 1350-1850), one of the largest Iron Age settlements in northeastern Taiwan [@Chen2007]. Kiwulan experienced European colonial powers, including the Spanish and Dutch in the 17th century, followed by a large influx of Han Chinese migrats in the 19th century, making it an ideal case to explore Indigenous responses to foreign presence over time. Key research questions include: How did Indigenous resource use change before and after these foreign encounters, particularly during the Dutch period of large-scale trade? Can we identify evidence of Indigenous participation in the deer trade from faunal remains? If so, any subsistence strategies emerged in response to economic demands from Europeans and Chinese traders? To address these questions, I analyze faunal remains with a focus on vertebrate species from Kiwulan, as zooarchaeology is an useful approach to studying local consumption and exchange systems [@Defrance2009; @Crabtree1990]. Through this approach, this research highlights the active role of Indigenous communities at the periphery of colonial bases within regional trade networks.

# The European colonization period and regional trade network in Northern Taiwan
<!-- Spanish colonization and the power relations in particular spatial and temporal contexts-->
In 1626, the Spanish (Spanish East Indies) established their first colony in northern Taiwan, San Salvador at Heping Dao, follwed by Santo Domingo in Tamsui (@fig-KWL-map), to counter the threat of the VOC (Vereenigde Oost-Indische Compagnie; Dutch East India Company) in the Asian market. The VOC had established a colony in southern Taiwan two years earlier [@Borao2009, p.41; @Andrade2007]. Although the Spanish intended to build a colonial city, San Salvador remained in its early stages of development throughout their colonization, consisting primarily of a fortress and a church [@Borao2009, pp. 103-134; @Berrocal2018]. The area surrounding the fortress was inhabited by Indigenous groups, the Taparris and the Quimaurris, who had lived at the site before being expelled by the Spanish [@Borao2001, SIT:165, p.81]. The relations between the Spanish and Indigenous people may not have involved extensive colonial control or intercultural domination, as Spanish territorial control appears to have been limited to the fort. Instead, the Spanish relied on exchange of local goods and food surpluses with Indigenous communities based on historical documents [@Borao2001, SIT:170-171]. This is consistent with the archaeological evidence from San Salvador, which shows a higher proportion of Indigenous and Chinese artifacts, with relatively few European materials [@Berrocal2018], indicating cultural entanglement.

<!-- the Dutch domination, how and why colonial powers were able to exert control -->
In 1642, the Dutch defeated the Spanish and took control of San Salvador and Santo Domingo, expanding their colonial power to northern and eastern Taiwan until 1662. Compared to the Spanish, the Dutch was more strategic in consolidating their power by establishing political and economic policies, forming alliances with Indigenous communities, and conquering resistant ones [@Kang2016; @Andrade2007]. Examples of colonial institutions imposed on Indigenous people include taxation, a tribute system, a village-leasehold system, census investigations, annual gatherings (Dutch: Landdag), and Indigenous representative selection [@Kang2016; @Andrade1997; @Andrade2007]. Among those colonial policies, the tribute system (1642-1647) and the village-leasehold system introduced in 1644 had profound impacts on Indigenous economies and the environment. Under the tribute system, each household in allied communities was required to pay deer products or rice to the Dutch annually, with deerskins being a primary item. In 1646, the amount of deerskin collected from Indigenous communities in northern Taiwan was 7,000 pieces [@Andrade2007]. Moreover, to increase deerskin yields, the Dutch implemented a hunting license system (1636-1645), which permitted Han Chinese to hunt deer on Indigenous lands during specific periods [@Heyns2001]. From 1633 to 1661, the average annual export of deerskins from Taiwan reached 67,000 [@Koo2011; @Cheng2017]. These colonial activities likely impacted the economy of Indigenous peoples, because deer was an important resource in prehistoric Taiwan [@Chen2000animal].

<!--variation of colonial powers across regions and indirect colonization at Kavalan-->
Although Dutch power expanded across nearly the entire Taiwan in the mid-17th century, the extent of their their colonial dominance varied greatly by region. Colonial control over areas far from the Dutch fortresses was often difficult and unsuccessful. Yilan in northeastern Taiwan is an example of weak colonial control or indirect colonization for most of the time [@Kang2012]. Yilan is located around 70 kilometers from European colonies with mountains in between, which made it difficult to access by land and was primarily reached by sea. In Yilan, Kavalan people was the main Indigenous group since the Iron Age. Due to the strategic position of Yilan on the route to Torrobouan, a place believed to have gold mines, the Dutch were eager to bring Yilan under their control [@DZI2000; @Kang2016, p47]. In 1644, the Dutch defeated several Kavalan communities and incorporated them into their colonial territory [@Kang2016]. However, instead of directly engaging with the Kavalan people, the Dutch relied on the assistance of intermediaries, the Basay people, who were active in Northern Taiwan, to collect tributes [@Kang2003]. This approach of indirect control allowed the Kavalan to maintain a degree of autonomy and even negotiate tributes [@Kang2016].

<!-- the pre-existing trade network in norththern Taiwan-->
Before the arrival of Europeans, an inter-insular trade network had been established since the 15th century in northern Taiwan, involving Chinese merchants and Indigenous groups in Tamsui and Keelung [@Chen2005]. Chinese traders transported goods such as textiles, porcelains, beads, and metal items in exchange for raw materials from Taiwan, including gold, sulfur, rattan, and animal hides and meats [@Chen2005; @Andrade2007]. This regional trade network was further connected to maritime trade between Ming China and the Ryukyu Kingdom around 15-16th centuries and later integrated into European long-distance trade in the 17th century [@Chen2005; @Von2019; @Wang2007]. Yilan was part of this pre-existing trade network, where local Kavalan people exchanged rice and deerskins for foreign goods and handicrafts from the Basay people [@Kang2003]. This trade network intensified during the period of European presence, which is consistent with the large quantities of imported foreign goods found at archaeological sites across Yilan, indicating frequent trade activity in the 17th century [@TaiwanHistory2005]. This unique historical and spatial context makes Yilan an ideal location for exploring Indigenous responses to foreign presence and the dynamics of trade networks in a pericolonial context [@Acabado2017; @Kang2016, pp. 31-78; @Trabert2017].

# Archaeological Context at Kiwulan
<!--introduction of Kiwulan and the population-->
Kiwulan, located in northern Yilan (@fig-KWL-map), is a multi-component site, with its upper component spanning from the late Iron Age to the historical periods (AD 1350-1850) [@Chen2007]. A large-scale salvage excavation at Kiwulan uncovered a total of 262 excavation units (4 x 4 m), revealing abundant artifacts and numerous features, including middens, burials, ditches, and post-holes with wooden posts in situ. The excavation suggests that Kiwulan was a well-developed settlement with longhouses (approximately 4 x 8 m) arranged side by side, and burials located to the north of the residential area [@Chen2004]. Based on oral history and the similarities between Kavalan material culture recorded in historical documents and the archaeological assemblage, Kiwulan is identified as a Kavalan settlement [@Chen2004]. The Kavalan people were the main Indigenous group in Yilan during the late Iron Age and early modern period. According to a Dutch census from the mid-17th century, there were approximately 40 Kavalan settlements in Yilan, with a total population of about 10,000 people [@Nakamura1938]. Among these Indigenous settlements, Kiwulan ("Kibannoran" in Dutch sources) was the largest one with 160 households and 711 inhabitants. In the early 19th century, increasing Han Chinese migration gradually led to the relocation of the Kavalan people to Hualien in eastern Taiwan [@Zhan1995]. This is reflected in the dramatic decline of Kiwulan’s population to fewer than 100.

<!--archaeological materials-->
The most common artifact found at Kiwulan was locally made pottery with geometric surface patterns, a typical type of ceramic in Yilan [@TaiwanHistory2005]. In addition to pottery, a variety of foreign goods, such as porcelains, stonewares, glass beads, carnelian beads, and metal artifacts and ornaments, were largely found in both residential area and burials. Since there is no evidence of local metallurgy or bead production, these items were likely obtained through exchange. This suggests that Kiwulan maintained continuous participation in regional trade networks linked to northern Taiwan, which were further influenced by long-distance trade in the 17th century [@Wang2007]. Moreover, foreign goods such as beads were considered prestige items or used for ritual purposes, and previous studies suggest that their introduction may have led to increasing social differentiation [@Wang2020ornament; @Wang2021; @Cheng2008].

<!--archaeological remains-->
Abundant animal bones, shellfish, and plant remains were uncovered at Kiwulan, reflecting the diversity of natural resources in the plain and an environment close to estuary. The faunal remains include deer, suids, cattle, birds, fish, and shellfish, while plant remains primarily consist of rice, peach, pandanus, and cucurbits [@Chen2004]. Rice farming appears to have been practiced, but likly limited a household scale rather than as intensive agriculture, according to archaeological evidence and historical records [@Chen1963]. The shellfish remains include bivalves such as clams and oysters, and univalves from freshwater and brackish water environments. Based on the archaeological remains and local environment, it is likely that Kiwulan residents practiced a mixed subsistence strategy, including rice farming, hunting, and fishing. Artifacts associated with these subsistence practices were also found, such as iron arrowheads for hunting, iron sickles for harvesting, and wooden paddles for water-related activities. Further research is still needed to examine resource use across different time periods and to better understand how subsistence strategies responded to foreign influences and environmental changes.

```{r}
#| label: fig-KWL-map
#| echo: false
#| fig-cap: "The geographic location of Kiwulan and European fortress in northern Taiwan (adapted from Wang and Marwick, 2020). Map data from naturalearthdata.com."

knitr::include_graphics(here::here("analysis/figures/KWL_map.png"))
```

# Materials and Methods
The faunal assemblages from Kiwulan analyzed in this study were sampled from 40 adjacent units at the residential area excavated during the first season (2001-2003) [@Chen2007]. These 40 units were selected because of their location that were believed to be the major residential area with thick archaeological deposits and relatively intact archaeological contexts without disruption from modern construction. The spatial continuity can provides a consistent framework for understanding food practices over time on a settlement scale. Also, the same excavation and recovery methods applied at the site allow comparison between assemblages from different temporal sequences. I mainly focus on vertebrate species due to their economic values in the trade market. The contexts of faunal remains include dwelling spaces and middens that were associated with daily dietary practices, whereas those from burials that might relate to ritual activities are excluded.

To explore changes in faunal remains over time, the original arbitrary levels of Kiwulan were assigned into six natural strata spanning from the 14th century to the 19th century, referred as cultural layers (CL1-Cl6) in this study. Each stratum was determined according to documentation of excavation depth, stratigraphic sections, sediment color and texture, radiocarbon dates, and archaeological diagnostic materials that served as time indicators. The diagnostic materials include stoneware commonly found at European associated sites in early 17th-century Southeast Asia, such as Anping jars and martavan [@Sakai2012; @Ketel2011; @Dizon2016; @Hsieh1995; @Lee2021], pipes, and Chinese bricks relating to the 19th century [@Wang2011; @Hsieh2009]. Although precisely identifying the stratum representing the specific time of European arrival is challenging due to its short duration, the separation of strata into six layers based on the criteria mentioned above could still offer a relatively robust framework for comparing archaeological records before and after the foreign influence. Among these six temporal layers, the fourth cultural layer (CL4) can be associated with the 17th century, when the Europeans occupied northern Taiwan, while the second cultural layer (CL2) corresponds to the 19th century, associated with the large influx of Chinese migrants in Yilan.

Only the faunal assemblages identified to at least the family level were considered in this study for consistent comparison across time sequences. Deer antlers without attaching to cranium were excluded since they were less likely related to dietary practices. To obtain generalized pattern of animal use, measures of ubiquity and taxonomic abundance were applied. Ubiquity measures the frequency of specific taxa through their presence or absence in particular depositional contexts, suggesting the regularity of consumtion of that taxa [@Mckechnie2016; @Vanderwarker2010]. Here ubiquity values are expressed as the percentage of units with a given taxon to the total number of units per cultural layers. Faunal assemblages were primarily quantified using the Number of Identified Specimens (NISP) method, with the Minimum Number of Individuals (MNI) applied when necessary, such as for estimating the expected NISP at a site [@Grayson2004; @Lyman2012]. Different from MNI that determines the actual numbers of animals recovered from remains, NISP offers a broad overview of remains by considering all fragmented bone assemblages. Despite the criticism of overestimation the number of animals, NISP captures better variation in species abundance on a site scale with less impacts from sample size [@Grayson1984; @Lyman2008]. For comparison across animal classes, the normalized NISP were used to avoid potential biases in identification due to differences in the preservation or recovery of skeletal elements [@Grayson2004]. The normalized NISP is calculated by dividing NISP by the number of skeletal elements.

Given that deerskin was a primary commodity in local regional exchange and in high demand during the Dutch colonization period [@Cheng2017], the frequencies of skeletal parts and surface modifications of deer bones were examined to assess butchery and transport strategies. This can shed light on how indigenous people utilize deer in response to high economic demand when involved in regional trade network. Skeletal parts were quantified using NISP and classified into groups based on meat yields and portions: meaty (e.g., upperlimb), mid-meaty (e.g., vertebrae), and less meaty parts (e.g., skull and lower limb) [@Madrigal2004; @Munro2004]. Teeth without associated cranial elements were excluded to reduce overrepresentation due to their much higher preservation rate and quantities compared to other parts. While lower limbs and skull bones have minimal meat, they are often valued in many cultures for marrow and soft tissues, such as tongue and brain [@Madrigal2002]. Thus, less meaty elements were considered as additional source of nutrition in the interpretation. Similarly, the degree of fragmentation could also imply the extraction of marrow or grease [@Lupo1997], which is calculated by dividing the Minimal Number of Elements (MNE) by NISP.

Deer bone surface modifications relative to human butchering and cooking practices, such as cut marks, fracturing, and burning damage, were recorded, as well as natural marks such as weathering and root etching for understanding taphonomic effects. Cut marks on major limb joints, such as shoulder, elbow, wrist, hip, knee, and ankle [@Lyman2005], were quantified to explore butchering patterns over time. Moreover, cut marks on lower limbs relative to skinning processes were examined and compared. Those marks are characteristic of circular and straight incisions left by the series of cuts for acquiring intact animal hides [@Costamagno2019; @Isaakidou2007].

The data and R code used for these analyses and visualizations are openly available at xxx for research reproducibility and transparency (@Marwick2017; @MarwickandWang2022).

# Results

```{r setup, include= FALSE}
library(tidyverse)
library(knitr)
library(viridis)
library(cowplot)
```

```{r}
#| label: get-data
#| echo: false

# read data
kwl_fauna <- readxl::read_excel(here::here("analysis","data","raw_data","KWL_faunal_P.xlsx"), sheet = 1,
                                col_types = "text")

kwl_fauna_HM <- readxl::read_excel(here::here("analysis","data","raw_data","KWL_faunal_HM.xlsx"), sheet = 1,
                                   col_types = "text")

kwl_fauna_broken <- readxl::read_excel(here::here("analysis","data","raw_data","broken_fauna.xlsx"), sheet = 1,
                                   col_types = "text")

sample_p <- readr::read_csv(here::here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv"))

kwl_chro_6 <- readxl::read_excel(here::here("analysis", "data", "raw_data", "KWL_chronology.xlsx"), sheet = 5)

# tidy chronological data
kwl_chro_tidy_6 <-
  kwl_chro_6 %>%
  mutate(area = str_sub(Pit, 5, 7), Pit = str_sub(Pit, 1, 4)) %>%
  mutate(area = ifelse(area == "", "ABCD", area)) %>%
  mutate(area = sapply(strsplit(area, ""), paste, collapse= ",")) %>%
  separate_rows(area) %>%
  pivot_longer(cols = starts_with("L"), names_to = "layer", values_to = "period") %>%
  mutate(layer = str_sub(layer, 2, 3)) %>%
  mutate(layer = ifelse(nchar(layer) == 1, paste0("0", layer), layer))

# combine faunal data of cultural layers and temporal sequences
fauna_sam <-
  kwl_fauna %>%
  mutate(Pit = paste(context, pit, sep = "")) %>%
  filter(Pit %in% sample_p$the_sq) %>%
  filter(componant == "上文化層" &!layer == "00") %>% #remove sc
  #filter(!str_detect(`總報告類別`, "[[:punct:]]+|[0-9]+")) %>%
  mutate(area = ifelse(is.na(area), "C", area)) %>% # assign an area to the blanks, which won't change the results
  left_join(kwl_chro_tidy_6, by = c("Pit" = "Pit", "layer" = "layer", "area" = "area"))

# combine faunal data of middens and temporal sequences
fauna_H_sam <-
  kwl_fauna_HM %>%
  select(-`空間號`, -`層   位`) %>%
  mutate(Pit = paste("P", pit, sep = "")) %>%
  filter(Pit %in% sample_p$the_sq) %>%
  filter(componant == "上文化層") %>%
  left_join(kwl_chro_tidy_6, by = c("Pit" = "Pit", "layer" = "layer", "area" = "area"))

fauna_combined_context_all <-
  rbind(fauna_sam, fauna_H_sam) %>% # combine the two datasets
  filter(!`部位/名稱` %in% c("角","犄角","角?")) %>%  #not associated with diet
  filter(!`部位/名稱`== "角基部"|!is.na(`部位/左右`)) %>% # associated with cranial parts
  mutate(`重量(g)` = as.numeric(`重量(g)`)) %>%
  group_by(taxa) %>%
  mutate(`Weight (g)` = sum(`重量(g)`, na.rm = T)) %>%
  ungroup() %>%
  mutate(period = factor(period, levels = c("CL1","CL2","CL3","CL4","CL5","CL6"), order = T)) %>%
  mutate(category = case_when(str_detect(animal, "sika")|str_detect(animal, "sambar")|
                              str_detect(animal, "e deer")~ "deer", str_detect(animal, "pig")|
                                str_detect(animal, "boar") ~ "suid", str_detect(animal, "fish") ~ "fish",
                              str_detect(animal, "cattle")|str_detect(animal, "buffalo")~ "bovine",
                              animal == "aves" ~ "bird", TRUE ~ animal)) %>%
  mutate(class = case_when(str_detect(taxa, "muntjac")|str_detect(taxa, "ammal")|str_detect(taxa, "erv")|
                           str_detect(taxa, "Sus")|str_detect(taxa, "Rattus")|str_detect(taxa, "Rusa")|
                           str_detect(taxa, "Bos")|str_detect(taxa, "Bubalus")~ "mammal",
                           str_detect(category, "fish") ~ "fish", str_detect(category, "bird") ~ "bird",
                           TRUE ~ "reptile")) %>%
  mutate(`部位/左右` = case_when(`部位/左右` == "Ｌ" ~ "L", `部位/左右` == "Ｒ" ~ "R", TRUE ~ `部位/左右`)) %>%
  mutate(category = fct_relevel(category, "deer","suid","muntjac","bovine","rat","bird","fish","turtle")) %>%
  mutate(cutmarks = ifelse(str_detect(`人為痕跡`, "切"), "yes", "no"),
         fractures = ifelse(str_detect(`人為痕跡`, "削")|str_detect(`人為痕跡`, "砍")|
                            str_detect(`人為痕跡`, "折")|str_detect(`人為痕跡`, "敲"), "yes", "no"),
         burning =  ifelse(str_detect(`人為痕跡`, "火")|str_detect(`人為痕跡`, "燒"), "yes", "no")) %>%
  filter(is.na(refitted)) # for NISP

fauna_combined_context <- fauna_combined_context_all %>% filter(!taxa == "Rattus sp.")
```

```{r}
#| label: taxa-NISP-counts
#| echo: false
# taxa counts
fauna_combined_taxa <-
  fauna_combined_context_all %>%
  count(taxa, `Weight (g)`) %>%
  mutate(taxa = str_replace(taxa, "cervidae", "cervid")) %>%
  mutate(taxa = case_when(str_detect(taxa, "mm") ~ paste("Indeterminate", tolower(taxa), sep =" "),
                          TRUE ~ taxa)) %>%
  mutate(`Common name` = case_when(str_detect(taxa, "Cer") ~ "Sika deer", str_detect(taxa, "Bub") ~ 
                                   "Water buffalo", str_detect(taxa, "Rus") ~ "Sambar deer",
                                   str_detect(taxa, "Ree") ~ "Muntjac", taxa == "Sus scrofa" ~ "Suid",
                                   str_detect(taxa, "Bos") ~ "Yellow cattle", str_detect(taxa, "Rat") ~ 
                                   "Rat", str_detect(taxa, "Pli") ~ "Catfish", str_detect(taxa, "Act") ~ 
                                   "Ray-finned fish", str_detect(taxa, "Tes") ~ "Turtle",
                                   str_detect(taxa, "cervid") ~ "Deer", taxa == "Sus scrofa domestica" ~ 
                                   "Domestic pig", taxa == "Aves" ~ "Birds")) %>%
  mutate(group = case_when(str_detect(`Common name`, "deer") ~ 1, str_detect(`Common name`, "Deer") ~ 2,
                           str_detect(taxa, "Sus") ~ 3, str_detect(taxa, "Bub")|str_detect(taxa, "Bos") ~ 4,
                           str_detect(taxa, "munt") ~ 5, str_detect(taxa, "Rat") ~ 6, 
                           str_detect(`Common name`, "fish") ~ 7, str_detect(taxa, "Tes") ~ 8,
                           str_detect(taxa, "Ave") ~ 9, str_detect(taxa, "large m") ~ 10,
                           str_detect(taxa, "medium m") ~ 11, TRUE ~ 12)) %>%
  group_by(group) %>%
  arrange(desc(`Weight (g)`), .by_group = TRUE) %>%
  ungroup () %>%
  rename(Taxon = taxa, NISP = n) %>%
  select(Taxon, `Common name`,  NISP, `Weight (g)`) %>%
  bind_rows(summarise(., across(where(is.numeric), sum), across(where(is.character), ~"Total"))) %>% 
  mutate(`Common name`= ifelse(is.na(`Common name`)|`Common name` == "Total", "-", `Common name`)) %>% 
  mutate(Taxon = ifelse(Taxon == "Sus scrofa", "Sus spp.", Taxon))

total_NISP <- fauna_combined_taxa %>% select(NISP) %>% tail(1) #rev(fauna_combined_taxa$NISP)[1]
unident <- fauna_combined_taxa %>% filter(str_detect(Taxon, "In")) %>% summarise(sum(NISP))
deer_NISP <- fauna_combined_taxa %>% 
  filter(str_detect(`Common name`, "eer")) %>% summarise(sum(NISP)) #|str_detect(Taxon, "Ree")
suid_NISP <- fauna_combined_taxa %>% filter(str_detect(Taxon, "Sus spp.")) %>% summarise(sum(NISP))
buff_NISP <- fauna_combined_taxa %>% filter(str_detect(Taxon, "Bub")) %>% summarise(sum(NISP))
deer_NISP_per <- round(deer_NISP/(total_NISP-unident), 2)*100
suid_NISP_per <- round(suid_NISP/(total_NISP-unident), 2)*100
buff_NISP_per <- round(buff_NISP/(total_NISP-unident), 2)*100
```

The faunal assemblage analyzed in this study from Kiwulan is composed of `r total_NISP` bone fragments based on NISP (@tbl-counts). Among the identifiable bones, large deer is the majority (`r deer_NISP_per`%), followed by suid (`r suid_NISP_per`) and water buffalo (`r buff_NISP_per`). To further compare general patterns of animal consumption in terms of their economic values over time, animals in the same family with similar body size were grouped together. Thus, the deer category includes sika deer, sambar deer, and large deer, while water buffalo and yellow cattle were group into the category, bovine. Similarly, domestic pig (n = 2, from CL4 and CL6) was integrated into the suid category. Rats (Rattus sp.) were excluded from further analyses as they are normally considered commensal animals, and their relatively intact skeletons from one single archaeological context show no direct evidence of being part of the daily diet. In general, the faunal remains show the frequencies increased from CL6, peaked at CL4, and then declined dramatically after CL4 ( (@tbl-counts, @fig-NISP-proportion: A). A similar trend was also observed in the weight of unidentified small bone fragments (@fig-NISP-proportion: B).

```{r}
#| label: tbl-counts
#| echo: false
#| tbl-cap: "Taxa, Number of Identified Specimens (NISP), and weight of faunal remains from the sampled area at Kiwulan."
#| tbl-colwidths: [40,25,15,20]
#| tbl-cap-location: bottom
library(kableExtra)
knitr::kable(fauna_combined_taxa, align = "l") 
```

```{r}
#| label: mammal-counts
#| echo: false
mammal_count <-
  fauna_combined_context %>%
  filter(class == "mammal") %>%
  count(period, category) %>% 
  drop_na() %>% 
  bind_rows(summarise(., across(where(is.numeric), sum), .by = period)) %>% 
  mutate(category = as.character(category)) %>% 
  mutate(category = ifelse(is.na(category), "total", category))

CL4 <- mammal_count %>% filter(period == "CL4") 
CL4_boar <- round(CL4$n[3]/CL4$n[4], 3)*100

# taxonomic diversity
category_count <-
  fauna_combined_context %>%
  count(period, category) %>% 
  count(period, category) %>% 
  drop_na() %>% 
  bind_rows(summarise(., across(where(is.numeric), sum), .by = period)) %>% 
  slice(24:29) %>% 
  select(period, n) %>% 
  rename(diversity = n)

# wide format and join
wider_mammal_count <-
  mammal_count %>% 
  pivot_wider(id_cols = period, names_from = category, values_from = n, values_fill = 0) %>% 
  left_join(category_count)

boar_reg <- summary(lm(suid ~ total, data = wider_mammal_count))
boar_lm <- c(round(boar_reg$adj.r.squared, 2), sprintf('%.3f', boar_reg$coefficients[,4][2]))
munt_reg <- summary(lm(muntjac ~ total, data = wider_mammal_count))
munt_lm <- c(round(munt_reg$adj.r.squared, 2), sprintf('%.2f', munt_reg$coefficients[,4][2]))
deer_reg <- summary(lm(deer ~ total, data = wider_mammal_count))
div_reg <- summary(lm(diversity ~ total, data = wider_mammal_count))
div_lm <- c(round(div_reg$adj.r.squared, 2), sprintf('%.2f', div_reg$coefficients[,4][2]))
```

For animal classes, the relative abundance shows that mammals are the primary dietary source throughout all layers, with fish as the secondary source in CL4 and birds in CL3 and CL2 (@fig-NISP-proportion: C). Fish and birds are both identified from CL5 to CL3; however, their proportions (less than 10%) are much lower compared to mammals. The reptile is represented by one piece of turtle shell found in CL4, which might not be a common food content at the site. Within those identifiable mammals, deer is the main animal from CL6 to CL2 that accounts for at least 85% per layer, while CL1 is predominated only by bovine (@fig-NISP-proportion: D). Suid was present from CL6 to CL3, and had a higher proportion in CL4 (`r CL4_boar` %), while muntjac appeared in CL5, and was continuously present until CL2. In general, there is an increase in animal diversity in CL4 and CL5, especially in CL4, where suid and fish have slightly higher proportions compared to those in other layers. Given that the taxonomic richness, or animal diversity, could be impacted by sample size varied from layer to layer, the relationship between the number of taxa and sample size (NISP) was examined by linear regression model. The result shows no correlation between the diversity of animals and sample size ($R^2$ = `r div_lm[1]`, P = `r div_lm[2]`), which suggests little bias from potential sample size.

```{r}
#| label: fig-NISP-proportion
#| echo: false
#| fig-cap: "A: The number of identified specimen per cultural layers. The NISP data is derived from Table 1 . CL6 represents the earliest layer and CL1 is the youngest. B: The weight of the unidentified small bone fragments for each cultural layer. C: Relative abundance of vertebrate classes by cultural layers based on normed NISP. D: Relative abundance of identified mammals by cultural layers based on NISP."

knitr::include_graphics(here::here("analysis/figures/NISP_pro_by_layer.png"))
```

```{r}
#| label: tbl-ubiquity
#| echo: false
#| tbl-cap: "Ubiquity of faunal categories across the sampling area at Kiwulan."
#| tbl-cap-location: bottom
# calculate the number of units for each cultural layers
kwl_unit_per_period <-
  kwl_chro_6 %>%
  mutate(all = pmap_chr(select(., -Pit, -Grid), ~toString(unique(na.omit(c(...)))))) %>%  # unite and remove duplicates
  mutate(unit = str_extract(Pit, "[A-Z]+[0-9]{3}")) %>%
  distinct(unit, .keep_all = TRUE) %>%
  mutate(CL1 = ifelse(str_detect(all, "CL1"), unit, NA)) %>%
  mutate(CL2 = ifelse(str_detect(all, "CL2"), unit, NA)) %>%
  mutate(CL3 = ifelse(str_detect(all, "CL3"), unit, NA)) %>%
  mutate(CL4 = ifelse(str_detect(all, "CL4"), unit, NA)) %>%
  mutate(CL5 = ifelse(str_detect(all, "CL5"), unit, NA)) %>%
  mutate(CL6 = ifelse(str_detect(all, "CL6"), unit, NA))

# count the units by cultural layers for later join
unit_by_period <-
  data.frame(unit_count = colSums(!is.na(kwl_unit_per_period))) %>%
  slice_tail(n = 6) %>%
  tibble::rownames_to_column("period")

# ubiquity of taxa from the 40 units by temporal sequences
fauna_ubiquity <-
  fauna_combined_context %>%
  select(Pit, layer, category, `重量(g)`, period, taxa) %>%
  count(period, Pit, category) %>%
  count(period, category) %>%
  drop_na(period, category) %>%
  left_join(unit_by_period) %>% # join the unit counts for each period
  mutate(ubiquity = paste(round(n/unit_count *100, 2), "%")) %>%
  select(-n, -unit_count) %>%
  pivot_wider(names_from = period, values_from = ubiquity) %>%
  mutate(across(CL1:CL6, ~replace_na(.x, "-"))) %>% 
  arrange(category)

CL4_ubi_deer <- fauna_ubiquity$CL4[1]
CL4_ubi_boar <- fauna_ubiquity$CL4[2]
CL4_ubi_munt <- fauna_ubiquity$CL4[3]

# get the bone mineral desity data
deer_element_BMD <- readr::read_csv(here::here("analysis", "data", "derived_data", "deer_BMD.csv"))
cor_sur_BMD <- cor.test(deer_element_BMD$NISP, deer_element_BMD$BMD)
survivor_cor <- round(c(cor_sur_BMD$estimate, cor_sur_BMD$p.value),2)

# for deer
fauna_deer_only <-
  fauna_combined_context %>%
  filter(category %in% c("deer") & !is.na(period))

deer_NISP <- fauna_deer_only %>% count(period) %>% rename(NISP = n)

# calculate MNE
deer_MNE <-
  fauna_deer_only  %>%
  filter(!is.na(`部位/左右`)&is.na(overlap)) %>% # remove overlapped fragments
  count(period,`部位/左右`,`部位/名稱`,`部位/位置`) %>%
  group_by(period) %>%
  summarise(MNE = sum(n), across())

deer_NISP <- fauna_deer_only %>% count(period) %>% rename(NISP = n)
frag_index <- left_join(deer_MNE, deer_NISP) %>% mutate(index = MNE/NISP) %>% distinct(period, index) %>% 
  mutate(index = round(index, 2))
frag_CL3_CL4 <- c(frag_index$index[2], frag_index$index[3])

knitr::kable(fauna_ubiquity)
```

@tbl-ubiquity shows deer had been the most widespread animal from CL6 to CL2, with the highest ubiquity in CL4 and CL5 (`r CL4_ubi_deer`), and lowest in CL2. The second and third animal regularly used are suid and muntjac. Although those two animals reached their highest ubiquity in CL4 (`r CL4_ubi_boar` for suid, and `r CL4_ubi_munt` for muntjac), their presence patterns are slightly different. Suid was utilized as early as deer in CL6, declined in CL3, and disappeared in CL2, whereas muntjac showed a cultural lag, appearing one phase later than suid. Bovine had a low ubiquity in CL5 and high ubiquity in CL6. Small animals (bird, fish, and turtle) had much lower ubiquity than large animals and they had a slightly increase in ubiquity in CL3. The result of animal ubiquity correspond with their frequencies (NISP), suggesting that deer is the most common resource at the site, with an increase utilization of suid and muntjac in CL4.

Although deer was predominant throughout all temporal sequences, variations in their skeletal parts could be identified. Deer bones were grouped into six body portions [@Madrigal2004], including head (cranium, mandible), axial (vertebrae, sternum, ribs), upper forelimb (scapula, humerus, ulna, and radius), upper hindlimb (femur, tibia, innominate), lower limb (carpals, metacarpals, tarsals, metatarsals, astragalus, calcaneus), and foot (phalanges). The result (@fig-deer-portion: A) shows that the meaty portion, especially the upper forelimb, is the most frequent part from CL6 to CL2, followed by upper hindlimb and the less-meaty portions, lower limb and foot. The consistency of higher numbers of upper limb and lower limb suggests that those parts should be transported together [e.g., @Stiner1991]. The generally low frequency of axial elements might be likely due to a high degree of fragmentation, butchery practices, or dietary preferences. A different pattern between layers is observed in the frequency of cranial bones, for which a few present in CL6, increased to the highest number in CL4, and then decreased thereafter. In addition, lower limb and foot bones appear to be proportionally less in CL2 and CL3 compared to earlier periods. Those differences might indicate changes in processing decisions of meat portions of deer.

Considering the impacts of differential bone density on preservation of skeletal elements, the relationship between the survivorship of deer skeletal elements and bone mineral density was tested by Pearson correlation. The values of bone survivorship were collected from @Lyman1994's research on deer, and calculated by dividing NISP by expected NISP, followed by multiplying 100%. The result (@fig-deer-portion: B) shows no significant positive correlation ($r$ = `r survivor_cor[1]`, P = `r survivor_cor[2]`)) between bone survivorship and bone density, suggesting that density-mediated bias has little impact on the taphonomic processes of bones. For the degree of bone fragmentation, CL3 shows higher fragmentation (index = `r frag_CL3_CL4[1]`), while bones in CL4 are less fragmented (index = `r frag_CL3_CL4[2]`). However, the fragmentation rates are not significantly differed between layers.

```{r}
#| label: fig-deer-portion
#| echo: false
#| fig-cap: "A: The frequency of skeletal parts of deer using NISP across cultural layers at Kiwulan. B: Bone minerial density agaist %survivorship of deer remains."

knitr::include_graphics(here::here("analysis/figures/deer_elements.png"))
```

```{r}
#| label: tbl-deer-modify
#| echo: false
#| tbl-cap: "Number and proportion of deer bones with surface modifications at Kiwulan."
#| tbl-cap-location: bottom
#| tbl-colwidths: [10,20,20,20,20,10]

# bones with surface modifications
burned_deer <- fauna_deer_only %>% filter(burning == "yes") %>% count(period) %>% rename(burned = n)
cutmark_deer <- fauna_deer_only %>% filter(cutmarks == "yes") %>% count(period) %>% rename(cutmarks = n)
frac_deer <- fauna_deer_only %>% filter(fractures  == "yes") %>%
  filter(`部位/名稱` %in% c("脛骨","股骨","肱骨","橈骨","上肢骨","掌骨","掌骨或蹠骨","蹠骨")) %>%
  count(period) %>% rename(fractures = n)
post_deposit <-fauna_deer_only %>% filter(!is.na(`保存狀態`)) %>% count(period) %>%
  rename(`post-deposition` = n)

list_modif <- list(burned_deer, cutmark_deer, frac_deer, post_deposit, deer_NISP)

# join
deer_modify <-
  list_modif %>%
  reduce(inner_join, by = 'period') %>%
  mutate(cut_per = round(cutmarks/NISP, 3) * 100, burn_per = round(burned/NISP, 3) * 100,
         frac_per = round(fractures/NISP, 3) * 100, post_per= round(`post-deposition`/NISP, 3) * 100) %>%
  mutate(burned = paste(burned, paste0("(", burn_per, " %)")),
         cutmarks = paste(cutmarks, paste0("(", cut_per, " %)")),
         fractures = paste(fractures, paste0("(", frac_per, " %)")),
         `post-deposition` = paste(`post-deposition`, paste0("(", post_per, " %)"))) %>%
  rename(`cut marks` = cutmarks) %>%
  select(-cut_per, -frac_per, -burn_per, -post_per)

knitr::kable(deer_modify)
```

Surface modifications, including cut marks, burning damage, and fractures were identified on deer bones from CL6 to CL2, with variations in their proportions (@tbl-deer-modify). Cut marks were present on 50% of bones in CL2, 30% of bones in earlier period (CL6 to CL4), and at a lower proportion of 23% in CL3. Most cut marks were located on major limb joints and long bones, suggesting butchering activities related to disarticulation and meat detaching. @fig-deer-cutmark shows that at least 50 % of joints in CL4 present cut marks, with a higher proportion of 75% at the hip. A similar pattern was observed in CL5, but with a lower proportion of cut marks on the ankle, knee, and hip compared to CL4. Not all joints were found in CL6, CL3 and CL2, making direct comparisons across all layers difficult. Nevertheless, the overall high proportion of joints with cut marks suggests intensive processing of deer throughout the site's occupation. In addition, cut marks on lower limbs should be associated with skinning (@fig-cutmark-img: e) and were primarily found on metapodials in CL6 (n = 1), CL5 (n = 4), CL4 (n = 6), CL3 (n = 1), followed by first phalanges in CL4 (n = 1) and CL5 (n = 1). Some joints displayed clean, smooth cut surfaces from chopping (@fig-cutmark-img: a, c), along with narrow V-shaped cut marks observed on most bones, indicating the use of metal tools [@Greenfield1999].

Burned bones (@fig-cutmark-img: d), characterized by charring and color changes, had the highest proportion in CL6 at 40% and gradually decreased to less than 10% after CL4. Bone breakage marks, including spiral fractures, transverse fractures, longitudinal splits, and percurssion scars [@Haynes1983; @Gifford2018; @Fisher1995], were observed frequently on long bones across different layers (@fig-fracture-img). These fractures indicate intentional marrow extraction at Kiwulan. CL2 had the highest proportion of fractured bones (> 50%), compared to 35% in CL3 and CL6, and 25% in CL4 and CL5. Chewing marks were observed on a few bones represented by a series of superficial elongated scores and circular depressions (@fig-fracture-img: e)[@Bello2023], suggesting carnivore gnawing. In addition, natural alterations to bones resulting from post-deposition impacts were also observed, such as weathering and root etching [@Gifford2018natural], on approximately one-third of the bones across all layers (@fig-fracture-img: c,d; @tbl-deer-modify). Those post-depositional impacts might have led to an underestimation of cut marks in general.

```{r}
#| label: fig-deer-cutmark
#| echo: false
#| fig-cap: "The proportion of cut marks on major limb joints of deer across cultural layers. The method for clasifying and quantifiying limb joints is based on Lyman (2005)."

knitr::include_graphics(here::here("analysis/figures/deer_cutmarks.png"))
```

```{r}
#| label: fig-cutmark-img
#| echo: false
#| fig-cap: "Examples of deer bones with cut marks, chop surfaces, and burning damage. Cut marks are indicated by red squares, and chop surfaces are indicated by black arrows. Bone fragments are (a) proximal radius with cut marks and a chop surface, (b) distal humerus with cut marks, (c) acetabulum with chop surfaces, (d) charred astragulus, (e) metatarsal with cut marks."

knitr::include_graphics(here::here("analysis/figures/deer_cutmark_img.png"))
```

```{r}
#| label: fig-fracture-img
#| echo: false
#| fig-cap: "Examples of deer bones with fractures, natural modifications, and chewing marks. Bone fragments are (a) proximal metatarsal with a puncturing hole and longitudinal split, (b) distal tibia showing spiral fracture, (c) distal radius showing tranverse fracture and root etching, (d) femur shaft showing longitudinal split, weathering, and a percurssion scar indicated by an arrow, (e) scapula with chewing marks."

knitr::include_graphics(here::here("analysis/figures/deer_fracture_img.png"))
```

# Discussion
<!--results and chronology -->
The faunal assemblage from the major residential area at Kiwulan suggests a continuous reliance on sika deer from CL6 to CL2, followed by a shift to bovine in CL1. Despite deer were the main animal resource during all times, variations in quantity, ubiquity, butchering patterns, and proportions relative to other animals suggest changes relative to foreign presence. The cultural layers were grouped into four temporal sequences in relation to foreign presences for the following interpretation. The sequences include the pre-European period (CL6, CL5), the European period (CL4), the post-European period (CL3), and the Chinese period with large Han Chinese settled down (CL2, CL1). The therm "European period" does not imply direct colonial control over Kiwulan but rather the broad impact of European colonial activities in northern Taiwan, such as long-distance trade, tribute systems, and occasional attacks affecting Indigenous communities in Yilan. Although the small sample size across periods limits comprehensive interpretation, examining results in a continuous sequence still reveals distinct patterns in animal use.

<!--summary of overall observations across time periods -->
The sampled area at Kiwulan contains the large quantity of faunal remains from the European period, followed by the late pre-European period. Interestingly, deer bones from the late pre-European period presented the same ubiquity as in the European period but only half the quantity. This might indicate a consistent pattern of regular deer consumption at the site both before and after European presence, with intensified deer use during the European period. In addition to the dominance of deer, the increased relative proportion and ubiquity of suids in the European period suggests a more regular use of suids compared to other periods. The dramatic decline in deer bones after the European period may be associated with both a reduction in the Kiwulan population and the decrease of wild deer. Sika deer populations in western Taiwan largely declined from the 18th century due to habitat loss and overhunting [@Chen2000animal]. A slight increase in muntjac use suggests that they may have served as a substitute, replacing suids as a secondary animal resource from earlier periods. Despite these shifts, deer remained a primary resource for local consumption at Kiwulan during the post-European and early Chinese periods. Bovines were present in small numbers (yellow cattle) in the pre-European period, but became the sole large animal (water buffalo) used at Kiwulan in the late Chinese period. Since bovines were absent from prehistoric sites in Taiwan and are widely believed to have been introduced from mainland China in historical period [@Chen2000animal], their increasing presence suggests Chinese influence and aligns with the Indigenous relocation to eastern Taiwan since the 19th century due to Chinese population pressures [@Zhan1995].

<!--skeletal part frequencies and butchery patterns of deer-->
Differences in deer skeletal completeness across periods suggests changes in butchery locations and cooking practices. There were more complete carcasses in the European and late pre-European periods compared to the post-European and Chinese periods. Moreover, the substantial rise in fragmented head parts (cranium and mandible) during the European period indicates increased use of soft tissues, such as brains and tongues, or the transport of more intact deer to the settlement for processing. This is consistent with the historical account in 1603, Record of Formosa (Dongfan ji) [@Chen1603], which describes the consumption of deer tongues and group hunting practices that could facilitate the transportation of deer. Changes in cooking practices are also evident. The decline in burned bones over time, with the lowest percentage in the European period, suggests a shift from roasting to other cooking methods, such as boiling [@Lupo1997]. The size of chopped bones, which could fit into local ceramics at Kiwulan, might be associated with boiling practices. This could be further confirmed through organic residue analysis on pottery. Other meat preparation methods, such as drying through salting and smoking, are also possible [e.g., @Spyrou2019]. Although vertebrae and ribs, typically used for drying [@Friesen2001; @Binford1978], were present in low frequencies at Kiwulan, historical records mentioning the consumption or exchange of dried meat by the Kavalan people supports the possibility [@Chen1963]. The shift in cooking practices might result from the increased need for deer, leading people to adopt boiling for its benefits, such as maximizing the use of residual meat, facilitating grease rendering, and improving quality of lean meat [@Lupo1997], while drying helped preserve and transport the meat.

A consistent pattern of high frequencies of upper limb bones but low frequencies of vertebrae and ribs was identified across all periods. This might suggest selective butchering and a preference for meaty portions. Moreover, among the meaty portions, front quarters (humerus and radius) were consistently more abundant than hindquarters (femur and tibia). One possible explanation is that femurs and tibiae were heavily fragmented due to marrow extraction, as these bones have high marrow yields [@Madrigal2002; @Lupo1998], making them less identifiable in the assemblage. Evidence for marrow extraction is supported by at least 25% of long bones showing fractures and a moderate degree of fragmentation overall. Another explanation is that some hindquarters were used for trade, as Dutch records mention that deer legs and meat were commodities in addition to hides [@Blusse2006, pp.220-224], but they did not specify which legs. If deer legs, particularly hindquarters, were traded as bone-meat packages, this could account for the lower frequencies of hindquarter bones compared to front quarters. While this hypothesis requires further evidence, the overall high proportion of meaty portions at Kiwulan suggests that deer meat might be not only used for local consumption but also processed for trade. Thus, both factors could contribute to this pattern, rather than a single factor alone.

<!--deer hides-->
Deerskin trade had been prevalent in East Asia since the 14th century and was monopolized by Europeans in the 17th century when a large number of hides, especially from western Taiwan, were exported to Japan [@Cheng2017; @Hollmann1991; @Wang1981]. The cut marks identified on deer lower limbs at Kiwulan confirm the removal of deer skins, likely for trade with external traders. Moreover, the abundance of foreign goods at Kiwulan, including ornaments, porcelains, and metal items, further supports the frequent exchanges between local residents and outsiders [@Wang2020ornament; @Chen2007]. This suggests that northeastern Taiwan, like western Taiwan, was actively involved in and impacted by long-distance trade. Although lower limbs with skinning cuts did not show a substantial increase during the European period, the disproportionate rise in head parts might hint at the likelihood of brain tanning when preparing hides. Brain tanning is a traditional method for softening and preserving hides in many Indigenous cultures, including those of North America, Southern Africa, North Asia, and Japan [@Baillargeon2010; @Badenhorst2009; @Hatt1969; @Matsumoto2016]. In Taiwan, the present-day Bunun group in the central highlands has revived hide tanning based on oral history, primarily using peanuts or pig lard as tanning materials, with occasional use of animal brains [@Li1999, pp. 48; @Xu1994]. Despite the specific tanning materials used by the Kavalan people remain unknown due to the loss of traditions, brain tanning is a plausible method given the availability of deer heads from hunting.

<!-- population dynamics and arguments-->
The high population during the European period, as inferred by the frequencies of pottery, may have contributed to the overall increase in animal use [@Wang2020pot; @Hsieh2009]. However, the impact of foreign presence can be still be identified in deer butchery patterns, the relative proportions of animals, and animal ubiquity. First, lower limbs with skinning cuts were mostly found in the European period, followed by the late pre-European phase, aligning with the rise of regional trade in the 14th–15th centuries and its intensification in the 17th century. Second, if the high frequency of deer heads during the European period was solely due to food stress from population growth, leading to increased brain or tongue consumption, a parallel rise in marrow extraction would be expected. However, long bones with fracture patterns showing 10–30% less frequent during the European and late pre-European periods than in other periods. Third, population growth would have likely led to a more proportional increase in both large and small animals. Despite high taxonomic diversity during the European period, suids showed a disproportionate rise in ubiquity, quantity, and proportion. This indicates that the trade of deerskin and meat may have led Indigenous people to rely more on suids as an alternative prey due to their comparable body size and meat return rates to deer [@Broughton2011prey]. The role of suids as a secondary resource after deer has been observed in Indigenous cultures in western Taiwan [@Chuang2021], suggesting they may have served as a supplement when deer were scarce or limited, possibly because some deer were reserved for trade.

After the Dutch left, the deerskin trade continued but on a much smaller scale, with around 10,000 pieces exported annually in the 18th century, which is far fewer than the 100,000 pieces exported during the European period [@Nakamura1997; @Tsao2011]. The shrinking trade market and declining population at Kiwulan may have contributed to the reduced reliance of suids, as shown in their lower proportion in the post-European period and eventual absence in the Chinese period. It should be noted that muntjac were consistently part of local subsistence, though in small numbers, until the early Chinese period. Their skins were used by many Indigenous groups such as the Tsou and Bunun for making clothing, similar to the use of deerskins [@Li1999], which may have also been practiced at Kiwulan. Although both suids and muntjacs served as substitutes for deer, their differing frequencies and ubiquity over time suggest that suids were the primary supplementary meat source, while muntjact were less frequently consumed but may have been valued for their hides in daily clothing.

<!--indigenous reaction and resilience-->
The Dutch played a dual role in Yilan as both colonizers demanding tribute, and traders integrated into the pre-existing inter-insular trade network involving Han Chinese merchants and different Indigenous groups [e.g., @Torrence2000]. That means, on one hand, Indigenous group in Yilan faced pressures from colonial encounters, such as occasional Dutch attacks, despite the generally indirect and weak colonial for most of the time. On the other hand, rather than resisting, local Indigenous people might participate volentarily in trade to acquire foreign necessities (e.g. metal knives and wares) and goods (e.g. ornaments and porcelains) that were highly valued in local culture [@Wang2020ornament]. The intensified hunting during the European period was not just a result of population growth but was driven by the deerskin trade. Zooarchaeological evidence from Kiwulan indicates that deer had been a primary food source and trade commodity since the late pre-European period, with its economic significance increasing during European presence. In addition, ethnographic studies suggest that deer hunting is closely tied to masculinity and coming-of-age ceremonies in Kavalan culture. That means, deer consumption at Kiwulan reflects not only daily practices but also the continuation of cultural traditions. Everyday practices play an essential role in maintaining Indigenous traditions within colonial settings and periods of frequent cultural exchange [@Silliman2009; @Noe2023]. In the pericolonial context of northeastern Taiwan, the continued consumption of deer, even as other animals like bovines were introduced in the pre-European period, highlights the persistence of cultural practices, while the increased reliance on suids at Kiwulan likely represents an adaptive and temporary strategy to meet high trade demands for deer products.

# Conclusion
By situating zooarchaeological evidence within the context of foreign presence, the Europeans in the 17th century and Chinese immigrants in the 19th century, this study reveals Indigenous participation in regional trade networks and their adaptation strategies. The frequencies, butchery patterns, and modifications on deer bones from Kiwulan indicates an active deer processing that was intensified during the European period. Deer as a primary animal resource at Kiwulan suggests cultural continuity, while the strategic reliance on suids during the European period reflects adaptation to high trade demands for deer products. Additionally, the incorporation of muntjac into the diet during the Chinese period likely resulted from the overhunting of sika deer. These findings highlight how Kiwulan residents navigated foreign influence and environmental changes while maintaining cultural traditions. Through the long-term dynamics of animal use at Kiwulan, this research demonstrates that even Indigenous communities in peripheral areas with relatively weak colonial control were profoundly impacted by colonial activities than previously thought. Future research expanding to other Indigenous settlements in northeastern Taiwan, with additional dietary evidence such as botanical remains and organic residue analysis, could provide deeper insights into regional dietary patterns and human-environment interactions.

# Acknowledgements
I would like to thank the Yilan County Cultural Affairs Bureau in Taiwan for permitting access to the zooarchaeological remains used in the analyses and the archaeological team for the early documentation of faunal remains that provides a foundation for this research. Special thanks to Chien-Hsiang Lin at the Biodiversity Research Center, Academia Sinica, for his guidance and assistance in identifying fish bones. I also appreciate Jui-Yang Tsou for his valuable comments and assistance during the identification process. I would like to express my gratitude to Yin-Ru Chiang and his team at the Biodiversity Research Center, Academia Sinica, for providing me a lab space to work on data and finalizing this work. I am also grateful to the reviewers of this manuscript for their comments.

\newpage

# References

<!-- The following line ensures the references appear here for the MS Word or HTML output files, rather than right at the end of the document (this will not work for PDF files):  -->

::: {#refs}
:::

\newpage

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r}
#| label: colophon
#| cache: false

# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```